---
title: "Caretaker_Preprocessing"
author: "Aulia Dini Rafsanjani"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(readr)
```

Abbreviation :

- MAP = Mean Arterial Pressure

- HR = Heart Rate

Rs with data type 1:

Rs with data type 2:

### Read Dataset

```{r}
# Read dataset
vitals_12452 <- read_csv(
  "O:/WorkLife-Data/biomonitors/caretaker_cbp/12452/12452_vitals_2024-07-05_17-01-56-1.csv", 
  show_col_types = FALSE)

vitalsavg_12452 <- read_csv(
  "O:/WorkLife-Data/biomonitors/caretaker_cbp/12452/12452_vitalsavg_2024-07-05_17-01-56-1.csv", 
  show_col_types = FALSE)

secondvitals_12452 <- read_csv(
  "O:/WorkLife-Data/biomonitors/caretaker_cbp/12452/12452_secondaryVitals_2024-07-05_17-01-56-1.csv",
  show_col_types = FALSE)
```

### Check data structure

```{r}
# Check data structure
str(vitals_12452)
```
```{r}
str(vitalsavg_12452)
```

```{r}
str(secondvitals_12452)
```

### Compute time difference

```{r}
library(dplyr)

vitals_12452 <- vitals_12452 %>%
  mutate(Diff = vitals_12452$`TimeStamp (mS)` - lag(vitals_12452$`TimeStamp (mS)`))
```

```{r}
# IBI
head(vitals_12452$Diff)
summary(vitals_12452$Diff)
```
### Compute the HR in milliseconds

```{r}
vitals_12452 <- vitals_12452 %>%
  mutate(HR_milliseconds = if_else(`HeartRate (bpm)` == 0, NA_real_, 60 / `HeartRate (bpm)`*1000))

summary(vitals_12452$HR_milliseconds)
```
Findings:

- The time stamps interval is not equal with the HR in milliseconds. 

### Summary of datasets 1

### vitals

```{r}
summary(vitals_12452$`Systolic (mmHg)`)
summary(vitals_12452$`Diastolic (mmHg)`)
summary(vitals_12452$`MAP (mmHg)`)
summary(vitals_12452$`HeartRate (bpm)`)
summary(vitals_12452$`Respiration (Bpm)`)
summary(vitals_12452$AS)
summary(vitals_12452$SQE)
summary(vitals_12452$`TimeStamp (mS)`)
```
Finding:

- AS, SQE may not be meaningful information because the value is all zero and the vitals_avg does not present any summary from these two indicators. 

- Each row represents information per minute. 

- The time in the vitals is different from the time in the secondsvital. 

### vitals_avg

This dataset contains summary (min, max, mean, std dev) for the physiological indicator (Systolic, Diastolic, MAP, HR, RR) within five minutes interval. 

I assume that RR is Respiration Rate. 

I don't think we need to compute the summary of these indicator. 

### secondvitals

```{r}
summary(secondvitals_12452$Reserved1)
summary(secondvitals_12452$CO)
summary(secondvitals_12452$Reserved2)
summary(secondvitals_12452$LVET)
summary(secondvitals_12452$Reserved3)
summary(secondvitals_12452$CPO)
summary(secondvitals_12452$CPI)
summary(secondvitals_12452$PPV)
summary(secondvitals_12452$SVI)
summary(secondvitals_12452$SVR)
summary(secondvitals_12452$RPP)
summary(secondvitals_12452$SVV)
summary(secondvitals_12452$SV)
summary(secondvitals_12452$CI)
summary(secondvitals_12452$`TimeStamp (mS)`)
```
Findings:

- Variables with zero values: Reserved1, CO, LVET, CPO, CPI, SVI, SVR, RPP, SV, CI.

- Variables with values: Reserved2, Reserved3, PPV, SVV. PPV and SVV have the same values, maybe these values are not necessary?

- May need to find what Reserved2 and Reserved3 mean for?

### Read dataset type 2

```{r}
bbb_vitals_18682 <- read_csv(
  "O:/WorkLife-Data/biomonitors/caretaker_cbp/18682/bbb_vitals_2024-05-16_16-59-29.csv", 
  show_col_types = FALSE)

other_vitals_18682 <- read_csv(
  "O:/WorkLife-Data/biomonitors/caretaker_cbp/18682/other_vitals_2024-05-16_16-59-29.csv", 
  show_col_types = FALSE)
```

### Check data structure

```{r}
str(bbb_vitals_18682)
str(other_vitals_18682)
```

### Check summary

```{r}
summary(bbb_vitals_18682$Systolic)
summary(bbb_vitals_18682$Diastolic)
summary(bbb_vitals_18682$MAP)
summary(bbb_vitals_18682$HR)
summary(bbb_vitals_18682$'BV*')
summary(bbb_vitals_18682$CO)
summary(bbb_vitals_18682$IBI)
summary(bbb_vitals_18682$LVET)
summary(bbb_vitals_18682$SV)
summary(bbb_vitals_18682$'AS*')
```
```{r}
summary(other_vitals_18682$Respiration)
```

Findings:

- BV*, CO, LVET, SV, AS* has a zero values --> Maybe these information is unnecessary?

- Respiration was presented in the other files --> other_vitals. 

### Code for time conversion

The Caretaker data shows a discrepancy of time in compare to the R's lab visit. The link of the time difference is located in this link: https://docs.google.com/spreadsheets/d/14YQDx0JMQStQE70ZA-V493Ft1qB48HpsVk1z30R_kW8/edit?gid=0#gid=0. 

Since I don't get any pattern of the time difference across all participant, we need to input the `new_start_time` manually. 

```{r}
# Load necessary library
library(lubridate)

# Convert to POSIXct
vitals_12452$Time <- as.POSIXct(vitals_12452$Time, format="%H:%M:%S", tz="UTC")

# Calculate time differences in seconds
vitals_12452$time_diffs <- c(0, diff(vitals_12452$Time))

# New start time
new_start_time <- as.POSIXct("17:15:00", format="%H:%M:%S", tz="UTC")

# Create new times
vitals_12452$new_times <- new_start_time + cumsum(vitals_12452$time_diffs)

# Format new times to keep only the HH:MM:SS part
vitals_12452$new_times <- format(vitals_12452$new_times, "%H:%M:%S")

# Print original and new times
print(head(vitals_12452$Time))
print(head(vitals_12452$new_times))
```
### Check the range of heart rate

```{r}
library(data.table)
library(dplyr)

# Define directory base path
base_directory <- "O:/WorkLife-Data/biomonitors/caretaker_cbp"

# Define participant IDs
participant_ids <- c("12452", "13098", "19003", "21854", "38190", "39925", "40761", "43465", "54010", "59154", "59503", "68129", "86760")

result_df1 <- data.frame()
result_df2 <- data.frame()

for (participant_id in participant_ids) {
  
  # Construct the directory path for the participant
  participant_directory <- file.path(base_directory, participant_id)
  
  # List all files in the participant's directory that match the pattern "ID_vitals_" and are CSV files
  pattern <- paste0(participant_id, "_vitals_.*\\.csv")
  full_file_paths <- list.files(participant_directory, pattern = pattern, full.names = TRUE)
  
  for (full_file_path in full_file_paths) {
    # Read the CSV file
    data <- fread(full_file_path, sep = ",")
    
    # Convert heart rate from bpm to ms
    data$HR_ms <- round(((60 / data$`HeartRate (bpm)`) * 1000), 3)
    
    # Compute time stamp differencew
    data$Diff = data$`TimeStamp (mS)` - lag(data$`TimeStamp (mS)`)
    
    # Create table of RR summary (in milliseconds)
    table1 <- data %>%
      summarise(minimum = round(min(HR_ms, na.rm = TRUE), 2),
                q1 = round(quantile(HR_ms, 0.25, na.rm = TRUE), 2),
                median = round(median(HR_ms, na.rm = TRUE), 2),
                q3 = round(quantile(HR_ms, 0.75, na.rm = TRUE), 2),
                maximum = round(max(HR_ms, na.rm = TRUE), 2),
                mean = round(mean(HR_ms, na.rm = TRUE), 2),
                sd = round(sd(HR_ms, na.rm = TRUE), 2)) %>%
      mutate(data_name = participant_id)
    
    # Create table of RR summary (in milliseconds)
    table2 <- data %>%
      summarise(minimum = round(min(Diff, na.rm = TRUE), 2),
                q1 = round(quantile(Diff, 0.25, na.rm = TRUE), 2),
                median = round(median(Diff, na.rm = TRUE), 2),
                q3 = round(quantile(Diff, 0.75, na.rm = TRUE), 2),
                maximum = round(max(Diff, na.rm = TRUE), 2),
                mean = round(mean(Diff, na.rm = TRUE), 2),
                sd = round(sd(Diff, na.rm = TRUE), 2)) %>%
      mutate(data_name = participant_id)
    
    
    # Append the percentage table to the result data frame
    result_df1 <- bind_rows(result_df1, table1)
    result_df2 <- bind_rows(result_df2, table2)
  }
}

# Reorder columns if needed
# Print Result
result_df1 <- result_df1[, c("data_name", "minimum", "q1", "median", "q3", "maximum", "mean", "sd")]
result_df1

result_df2 <- result_df2[, c("data_name", "minimum", "q1", "median", "q3", "maximum", "mean", "sd")]
result_df2
```

Does the HR based on this calculation fall within the range of 0.3s - 1.3s or 300ms - 1300ms ? Yes, except some participant has Inf values because they have HR with 0 values. I use the formula of HR_ms = 60/HR_bpm

Does the HR based on this calculation fall within the range of 0.3s - 1.3s or 300ms - 1300ms ? No, if I use the calculation based on Timestamp difference. 

### Check the calculation of MAP

```{r}
vitals_12452$map_check = round((vitals_12452$`Diastolic (mmHg)` + ((1/3)*(vitals_12452$`Systolic (mmHg)` - vitals_12452$`Diastolic (mmHg)`))),0)

vitals_12452$diff_map = vitals_12452$map_check - vitals_12452$`MAP (mmHg)`
summary(vitals_12452$diff_map)
```
```{r}
hist(vitals_12452$diff_map)
```
### Check the IBI in bbb_vitals VS the difference of timestamp

```{r}
library(data.table)
library(dplyr)

# Define directory base path
base_directory <- "O:/WorkLife-Data/biomonitors/caretaker_cbp"

# Define participant IDs
participant_ids <- c("38190", "59503", "59154", "19599", "21854", "13098", "39925", "19686")

result_df1 <- data.frame()

for (participant_id in participant_ids) {
  
  # Construct the directory path for the participant
  participant_directory <- file.path(base_directory, participant_id)
  
  # List all files in the participant's directory that match the pattern "ID_vitals_" and are CSV files
  pattern <- paste0(participant_id, "_vitals_.*\\.csv")
  full_file_paths <- list.files(participant_directory, pattern = pattern, full.names = TRUE)
  
  for (full_file_path in full_file_paths) {
    
    # Read the CSV file
    data <- fread(full_file_path, sep = ",")
    
    # Compute time stamp differencew
    data$Diff = data$`TimeStamp (mS)` - lag(data$`TimeStamp (mS)`)
    
    # Create table of RR summary (in milliseconds)
    table1 <- data %>%
      summarise(min_diff = round(min(Diff, na.rm = FALSE), 2),
                q1_diff = round(quantile(Diff, 0.25, na.rm = TRUE), 2),
                median_diff = round(median(Diff, na.rm = TRUE), 2),
                q3_diff = round(quantile(Diff, 0.75, na.rm = TRUE), 2),
                max_diff = round(max(Diff, na.rm = TRUE), 2),
                mean_diff = round(mean(Diff, na.rm = TRUE), 2),
                sd_diff = round(sd(Diff, na.rm = TRUE), 2)) %>%
      mutate(data_name = participant_id)
    
    
    # Append the percentage table to the result data frame
    result_df1 <- bind_rows(result_df1, table1)
  }
}

# Reorder columns if needed
# Print Result
result_df1 <- result_df1[, c("data_name", "min_diff", "q1_diff", "median_diff", "q3_diff", "max_diff", "mean_diff", "sd_diff")]
result_df1
```

```{r}
library(data.table)
library(dplyr)

# Define directory base path
base_directory <- "O:/WorkLife-Data/biomonitors/caretaker_cbp"

# Define participant IDs
participant_ids <- c("38190", "59503", "59154", "19599", "21854", "13098", "39925", "19686")

result_df1 <- data.frame()

for (participant_id in participant_ids) {
  
  # Construct the directory path for the participant
  participant_directory <- file.path(base_directory, participant_id, "Type_2_Data")
  
  # List all files in the participant's directory that match the pattern "ID_vitals_" and are CSV files
  pattern <- paste0("bbb_.*\\.csv")
  full_file_paths <- list.files(participant_directory, pattern = pattern, full.names = TRUE)
  
  for (full_file_path in full_file_paths) {
    
    # Read the CSV file
    data <- fread(full_file_path, sep = ",")
    
    # Create table of RR summary (in milliseconds)
    table1 <- data %>%
      summarise(min_IBI = round(min(IBI, na.rm = TRUE), 2),
                q1_IBI = round(quantile(IBI, 0.25, na.rm = TRUE), 2),
                median_IBI = round(median(IBI, na.rm = TRUE), 2),
                q3_IBI = round(quantile(IBI, 0.75, na.rm = TRUE), 2),
                max_IBI = round(max(IBI, na.rm = TRUE), 2),
                mean_IBI = round(mean(IBI, na.rm = TRUE), 2),
                sd_IBI = round(sd(IBI, na.rm = TRUE), 2)) %>%
      mutate(data_name = participant_id)
    
    
    # Append the percentage table to the result data frame
    result_df1 <- bind_rows(result_df1, table1)
  }
}

# Reorder columns if needed
# Print Result
result_df1 <- result_df1[, c("data_name", "min_IBI", "q1_IBI", "median_IBI", "q3_IBI", "max_IBI", "mean_IBI", "sd_IBI")]
result_df1
```

Findings:

I am checking the summary statistics of the difference between time stamps with the IBI value. They have a different result. 

### Check the summary statistics of variable in data type 2

### 1. File name "bbb_vitals_.csv"

```{r}
library(data.table)
library(dplyr)

# Define directory base path
base_directory <- "O:/WorkLife-Data/biomonitors/caretaker_cbp"

# Define participant IDs
participant_ids <- c("38190", "59503", "59154", "19599", "21854", "13098", "39925", "19686")

result_df1 <- data.frame()

for (participant_id in participant_ids) {
  
  # Construct the directory path for the participant
  participant_directory <- file.path(base_directory, participant_id, "Type_2_Data")
  
  # List all files in the participant's directory that match the pattern "ID_vitals_" and are CSV files
  pattern <- paste0("bbb_.*\\.csv")
  full_file_paths <- list.files(participant_directory, pattern = pattern, full.names = TRUE)
  
  for (full_file_path in full_file_paths) {
    
    # Read the CSV file
    data <- fread(full_file_path, sep = ",")
    
    # Create table of RR summary (in milliseconds)
    table1 <- data %>%
      summarise(count_obs = nrow(data),
                min_syst = round(min(Systolic, na.rm = TRUE), 2),
                mean_syst = round(mean(Systolic, na.rm = TRUE), 2),
                max_syst = round(max(Systolic, na.rm = TRUE), 2),
                min_dias = round(min(Diastolic, na.rm = TRUE), 2),
                mean_dias = round(mean(Diastolic, na.rm = TRUE), 2),
                max_dias = round(max(Diastolic, na.rm = TRUE), 2),
                min_MAP = round(min(MAP, na.rm = TRUE), 2),
                mean_MAP = round(mean(MAP, na.rm = TRUE), 2),
                max_MAP = round(max(MAP, na.rm = TRUE), 2),
                min_HR = round(min(HR, na.rm = TRUE), 2),
                mean_HR = round(mean(HR, na.rm = TRUE), 2),
                max_HR = round(max(HR, na.rm = TRUE), 2),
                min_BV = round(min(`BV*`, na.rm = TRUE), 2),
                mean_BV = round(mean(`BV*`, na.rm = TRUE), 2),
                max_BV = round(max(`BV*`, na.rm = TRUE), 2),
                min_CO = round(min(CO, na.rm = TRUE), 2),
                mean_CO = round(mean(CO, na.rm = TRUE), 2),
                max_CO = round(max(CO, na.rm = TRUE), 2),
                min_IBI = round(min(IBI, na.rm = TRUE), 2),
                mean_IBI = round(mean(IBI, na.rm = TRUE), 2),
                max_IBI = round(max(IBI, na.rm = TRUE), 2),
                min_LVET = round(min(LVET, na.rm = TRUE), 2),
                mean_LVET = round(mean(LVET, na.rm = TRUE), 2),
                max_LVET = round(max(LVET, na.rm = TRUE), 2),
                min_SV = round(min(SV, na.rm = TRUE), 2),
                mean_SV = round(mean(SV, na.rm = TRUE), 2),
                max_SV = round(max(SV, na.rm = TRUE), 2),
                min_AS = round(min(`AS*`, na.rm = TRUE), 2),
                mean_AS = round(mean(`AS*`, na.rm = TRUE), 2),
                max_AS = round(max(`AS*`, na.rm = TRUE), 2)) %>%
      mutate(data_name = participant_id)
    
    # Append the percentage table to the result data frame
    result_df1 <- bind_rows(result_df1, table1)
  }
}

# Reorder columns if needed
# Print Result
result_df1 <- result_df1[, c("data_name", 
                             "count_obs",
                             "min_syst", "mean_syst", "max_syst",
                             "min_dias", "mean_dias", "max_dias",
                             "min_MAP", "mean_MAP", "max_MAP",
                             "min_HR", "mean_HR", "max_HR",
                             "min_BV", "mean_BV", "max_BV",
                             "min_CO", "mean_CO", "max_CO",
                             "min_IBI", "mean_IBI", "max_IBI",
                             "min_LVET", "mean_LVET", "max_LVET",
                             "min_SV", "mean_SV", "max_SV",
                             "min_AS", "mean_AS", "max_AS")]
result_df1
```

Finding:

- The variables with zero values : BV, CO, LVET, SV, AS

### 2. File name "other_vitals_.csv"

```{r}
library(data.table)
library(dplyr)

# Define directory base path
base_directory <- "O:/WorkLife-Data/biomonitors/caretaker_cbp"

# Define participant IDs
participant_ids <- c("38190", "59503", "59154", "19599", "21854", "13098", "39925", "19686")

result_df1 <- data.frame()

for (participant_id in participant_ids) {
  
  # Construct the directory path for the participant
  participant_directory <- file.path(base_directory, participant_id, "Type_2_Data")
  
  # List all files in the participant's directory that match the pattern "ID_vitals_" and are CSV files
  pattern <- paste0("other_vitals_.*\\.csv")
  full_file_paths <- list.files(participant_directory, pattern = pattern, full.names = TRUE)
  
  for (full_file_path in full_file_paths) {
    
    # Read the CSV file
    data <- fread(full_file_path, sep = ",")
    
    # Create table of RR summary (in milliseconds)
    table1 <- data %>%
      summarise(min_Respiration = round(min(Respiration, na.rm = TRUE), 2),
                mean_Respiration = round(mean(Respiration, na.rm = TRUE), 2),
                max_Respiration = round(max(Respiration, na.rm = TRUE), 2)) %>%
      mutate(data_name = participant_id)
    
    # Append the percentage table to the result data frame
    result_df1 <- bind_rows(result_df1, table1)
  }
}

# Reorder columns if needed
# Print Result
result_df1 <- result_df1[, c("data_name", 
                             "min_Respiration", "mean_Respiration", "max_Respiration")]
result_df1
```

### Check data type 1 

1. File "_vitals_.csv"

```{r}
library(data.table)
library(dplyr)

# Define directory base path
base_directory <- "O:/WorkLife-Data/biomonitors/caretaker_cbp"

# Define participant IDs
participant_ids <- c("38190", "59503", "59154", "19599", "21854", "13098", "39925", "19686")

result_df1 <- data.frame()

for (participant_id in participant_ids) {
  
  # Construct the directory path for the participant
  participant_directory <- file.path(base_directory, participant_id)
  
  # List all files in the participant's directory that match the pattern "ID_vitals_" and are CSV files
  pattern <- paste0("_vitals_.*\\.csv")
  full_file_paths <- list.files(participant_directory, pattern = pattern, full.names = TRUE)
  
  for (full_file_path in full_file_paths) {
    
    # Read the CSV file
    data <- fread(full_file_path, sep = ",")
    
    # Create table of RR summary (in milliseconds)
    table1 <- data %>%
      summarise(count_obs = nrow(data),
                min_syst = round(min(`Systolic (mmHg)`, na.rm = TRUE), 2),
                mean_syst = round(mean(`Systolic (mmHg)`, na.rm = TRUE), 2),
                max_syst = round(max(`Systolic (mmHg)`, na.rm = TRUE), 2),
                min_dias = round(min(`Diastolic (mmHg)`, na.rm = TRUE), 2),
                mean_dias = round(mean(`Diastolic (mmHg)`, na.rm = TRUE), 2),
                max_dias = round(max(`Diastolic (mmHg)`, na.rm = TRUE), 2),
                min_MAP = round(min(`MAP (mmHg)`, na.rm = TRUE), 2),
                mean_MAP = round(mean(`MAP (mmHg)`, na.rm = TRUE), 2),
                max_MAP = round(max(`MAP (mmHg)`, na.rm = TRUE), 2),
                min_HR = round(min(`HeartRate (bpm)`, na.rm = TRUE), 2),
                mean_HR = round(mean(`HeartRate (bpm)`, na.rm = TRUE), 2),
                max_HR = round(max(`HeartRate (bpm)`, na.rm = TRUE), 2),
                min_RR = round(min(`Respiration (Bpm)`, na.rm=TRUE), 2),
                mean_RR = round(mean(`Respiration (Bpm)`, na.rm=TRUE), 2), 
                max_RR = round(max(`Respiration (Bpm)`, na.rm=TRUE), 2),
                min_AS = round(min(AS, na.rm = TRUE), 2),
                mean_AS = round(mean(AS, na.rm = TRUE), 2),
                max_AS = round(max(AS, na.rm = TRUE), 2), 
                min_SQE = round(min(SQE, na.rm = TRUE), 2),
                mean_SQE = round(mean(SQE, na.rm = TRUE), 2),
                max_SQE = round(max(SQE, na.rm = TRUE), 2),
                min_TIME = round(min(`TimeStamp (mS)`, na.rm = TRUE), 2),
                mean_TIME = round(mean(`TimeStamp (mS)`, na.rm = TRUE), 2),
                max_TIME = round(max(`TimeStamp (mS)`, na.rm = TRUE), 2)) %>%
      mutate(data_name = participant_id)
    
    # Append the percentage table to the result data frame
    result_df1 <- bind_rows(result_df1, table1)
  }
}

# Reorder columns if needed
# Print Result
result_df1 <- result_df1[, c("data_name", 
                             "count_obs",
                             "min_syst", "mean_syst", "max_syst",
                             "min_dias", "mean_dias", "max_dias",
                             "min_MAP", "mean_MAP", "max_MAP",
                             "min_HR", "mean_HR", "max_HR",
                             "min_RR", "mean_RR", "max_RR",
                             "min_AS", "mean_AS", "max_AS",
                             "min_SQE", "mean_SQE", "max_SQE",
                             "min_TIME", "mean_TIME", "max_TIME")]
result_df1
```

### 2. File secondaryVitals

```{r}
library(data.table)
library(dplyr)

# Define directory base path
base_directory <- "O:/WorkLife-Data/biomonitors/caretaker_cbp"

# Define participant IDs
participant_ids <- c("38190", "59503", "59154", "19599", "21854", "13098", "39925", "19686")

result_df1 <- data.frame()

for (participant_id in participant_ids) {
  
  # Construct the directory path for the participant
  participant_directory <- file.path(base_directory, participant_id)
  
  # List all files in the participant's directory that match the pattern "ID_vitals_" and are CSV files
  pattern <- paste0("_secondaryVitals_.*\\.csv")
  full_file_paths <- list.files(participant_directory, pattern = pattern, full.names = TRUE)
  
  for (full_file_path in full_file_paths) {
    
    # Read the CSV file
    data <- fread(full_file_path, sep = ",")
    
    # Create table of RR summary (in milliseconds)
    table1 <- data %>%
      summarise(
                min_RESERVED1 = round(min(Reserved1, na.rm = TRUE), 2),
                mean_RESERVED1 = round(mean(Reserved1, na.rm = TRUE), 2),
                max_RESERVED1 = round(max(Reserved1, na.rm = TRUE), 2),
                min_CO = round(min(CO, na.rm = TRUE), 2),
                mean_CO = round(mean(CO, na.rm = TRUE), 2),
                max_CO = round(max(CO, na.rm = TRUE), 2),
                min_RESERVED2 = round(min(Reserved2, na.rm = TRUE), 2),
                mean_RESERVED2 = round(mean(Reserved2, na.rm = TRUE), 2),
                max_RESERVED2 = round(max(Reserved2, na.rm = TRUE), 2),
                min_LVET = round(min(LVET, na.rm = TRUE), 2),
                mean_LVET = round(mean(LVET, na.rm = TRUE), 2),
                max_LVET = round(max(LVET, na.rm = TRUE), 2),
                min_RESERVED3 = round(min(Reserved3, na.rm = TRUE), 2),
                mean_RESERVED3 = round(mean(Reserved3, na.rm = TRUE), 2),
                max_RESERVED3 = round(max(Reserved3, na.rm = TRUE), 2),   
                min_CPO = round(min(CPO, na.rm = TRUE), 2),
                mean_CPO = round(mean(CPO, na.rm = TRUE), 2),
                max_CPO = round(max(CPO, na.rm = TRUE), 2),
                min_CPI = round(min(CPI, na.rm = TRUE), 2),
                mean_CPI = round(mean(CPI, na.rm = TRUE), 2),
                max_CPI = round(max(CPI, na.rm = TRUE), 2), 
                min_PPV = round(min(PPV, na.rm = TRUE), 2),
                mean_PPV = round(mean(PPV, na.rm = TRUE), 2),
                max_PPV = round(max(PPV, na.rm = TRUE), 2),
                min_SVI = round(min(SVI, na.rm = TRUE), 2),
                mean_SVI = round(mean(SVI, na.rm = TRUE), 2),
                max_SVI = round(max(SVI, na.rm = TRUE), 2),
                min_SVR = round(min(SVR , na.rm = TRUE), 2),
                mean_SVR  = round(mean(SVR , na.rm = TRUE), 2),
                max_SVR  = round(max(SVR , na.rm = TRUE), 2),  
                min_RPP = round(min(RPP , na.rm = TRUE), 2),
                mean_RPP  = round(mean(RPP , na.rm = TRUE), 2),
                max_RPP  = round(max(RPP , na.rm = TRUE), 2),  
                min_SVV = round(min(SVV , na.rm = TRUE), 2),
                mean_SVV  = round(mean(SVV , na.rm = TRUE), 2),
                max_SVV  = round(max(SVV , na.rm = TRUE), 2),
                min_SV = round(min(SV , na.rm = TRUE), 2),
                mean_SV  = round(mean(SV , na.rm = TRUE), 2),
                max_SV  = round(max(SV , na.rm = TRUE), 2),
                min_CI = round(min(CI , na.rm = TRUE), 2),
                mean_CI  = round(mean(CI , na.rm = TRUE), 2),
                max_CI  = round(max(CI , na.rm = TRUE), 2),  
                min_TIMESTAMP = round(min(`TimeStamp (mS)` , na.rm = TRUE), 2),
                mean_TIMESTAMP = round(mean(`TimeStamp (mS)` , na.rm = TRUE), 2),
                max_TIMESTAMP = round(max(`TimeStamp (mS)` , na.rm = TRUE), 2)) %>%
      mutate(data_name = participant_id)
    
    # Append the percentage table to the result data frame
    result_df1 <- bind_rows(result_df1, table1)
  }
}

# Reorder columns if needed
# Print Result
result_df1 <- result_df1[, c("data_name", 
                             "min_RESERVED1", "mean_RESERVED1", "max_RESERVED1",
                             "min_CO", "mean_CO", "max_CO",
                             "min_RESERVED2", "mean_RESERVED2", "max_RESERVED2",
                             "min_LVET", "mean_LVET", "max_LVET",
                             "min_RESERVED3", "mean_RESERVED3", "max_RESERVED3",
                             "min_CPO", "mean_CPO", "max_CPO",
                             "min_CPI", "mean_CPI", "max_CPI",
                             "min_PPV", "mean_PPV", "max_PPV",
                             "min_SVI", "mean_SVI", "max_SVI",
                             "min_SVR", "mean_SVR", "max_SVR",
                             "min_RPP", "mean_RPP", "max_RPP",
                             "min_SVV", "mean_SVV", "max_SVV",
                             "min_SV", "mean_SV", "max_SV",
                             "min_CI", "mean_CI", "max_CI",
                             "min_TIMESTAMP", "mean_TIMESTAMP", "max_TIMESTAMP")]
result_df1
```


### 3. File "_vitalsavg_.csv"

```{r}
library(data.table)
library(dplyr)

# Define directory base path
base_directory <- "O:/WorkLife-Data/biomonitors/caretaker_cbp"

# Define participant IDs
participant_ids <- c("38190", "59503", "59154", "19599", "21854", "13098", "39925", "19686")

result_df1 <- data.frame()

for (participant_id in participant_ids) {
  
  # Construct the directory path for the participant
  participant_directory <- file.path(base_directory, participant_id)
  
  # List all files in the participant's directory that match the pattern "_vitalsavg_" and are CSV files
  pattern <- "_vitalsavg_.*\\.csv"
  full_file_paths <- list.files(participant_directory, pattern = pattern, full.names = TRUE)
  
  for (full_file_path in full_file_paths) {
    
    # Read the CSV file
    data <- fread(full_file_path, sep = ",")
    
    # Create table of summary statistics
    table1 <- data %>%
      summarise(count_obs = n(),
                mean_INTME = round(mean(`Interval Measurements`, na.rm = TRUE), 2),
                mean_SYSSTD = round(mean(`Sys StdDev`, na.rm = TRUE), 2),
                mean_SYSAVG = round(mean(`Sys Avg`, na.rm = TRUE), 2),
                mean_SYSMAX = round(mean(`Sys Max`, na.rm = TRUE), 2),
                mean_SYSMIN = round(mean(`Sys Min`, na.rm = TRUE), 2),
                mean_DIASTD = round(mean(`Dia StdDev`, na.rm = TRUE), 2),
                mean_DIAAVG = round(mean(`Dia Avg`, na.rm = TRUE), 2),
                mean_DIAMAX = round(mean(`Dia Max`, na.rm = TRUE), 2),
                mean_DIAMIN = round(mean(`Dia Min`, na.rm = TRUE), 2),
                mean_MAPSTD = round(mean(`MAP StdDev`, na.rm = TRUE), 2),
                mean_MAPAVG = round(mean(`MAP Avg`, na.rm = TRUE), 2),
                mean_MAPMAX = round(mean(`MAP Max`, na.rm = TRUE), 2),
                mean_MAPMIN = round(mean(`MAP Min`, na.rm = TRUE), 2),
                mean_HRSTD = round(mean(`HR StdDev`, na.rm = TRUE), 2),
                mean_HRAVG = round(mean(`HR Avg`, na.rm = TRUE), 2),
                mean_HRMAX = round(mean(`HR Max`, na.rm = TRUE), 2),
                mean_HRMIN = round(mean(`HR Min`, na.rm = TRUE), 2),
                mean_RRSTD = round(mean(`RR StdDev`, na.rm = TRUE), 2),
                mean_RRAVG = round(mean(`RR Avg`, na.rm = TRUE), 2),
                mean_RRMAX = round(mean(`RR Max`, na.rm = TRUE), 2),
                mean_RRMIN = round(mean(`RR Min`, na.rm = TRUE), 2)) %>%
      mutate(data_name = participant_id)
    
    # Append the summary table to the result data frame
    result_df1 <- bind_rows(result_df1, table1)
  }
}

# Reorder columns if needed
result_df1 <- result_df1[, c("data_name", 
                             "count_obs",
                             "mean_INTME", 
                             "mean_SYSSTD", "mean_SYSAVG", "mean_SYSMAX", "mean_SYSMIN",
                             "mean_DIASTD", "mean_DIAAVG", "mean_DIAMAX", "mean_DIAMIN", 
                             "mean_MAPSTD", "mean_MAPAVG", "mean_MAPMAX", "mean_MAPMIN", 
                             "mean_HRSTD", "mean_HRAVG", "mean_HRMAX", "mean_HRMIN", 
                             "mean_RRSTD", "mean_RRAVG", "mean_RRMAX", "mean_RRMIN")]

# Print the final result
result_df1

```
