---
title: "Caretaker_Filtering_Imputation"
author: "Aulia Dini Rafsanjani"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This file contains final code of two parts: Filtering and Imputation. The filtering process is handled by the `filtering` function, while the imputation process is managed by the `impute_IBI_intervals` function. To run the code, first execute the `filtering` code chunks and apply the filtering function to your data set. After filtering, you can use the resulting data_filtered as input for the `impute_IBI_intervals` steps.

Please note that the processing time may take a few minutes to complete. Therefore, I recommend running this code for each data set individually (rather than many data set simultaneously) to avoid potential errors.

To apply this code to real participants, you should use the new naming convention format in the O drive. 

Side note:

The imputation code use the random number from normal distribution to fill out the gap. I use the function `rnorm` without using `set.seed` function. This means that every time we do the imputation for the same data set, the result will generate a different random number. This code has commented out the set.seed() function. The code can be activated anytime once the team got a answer from Michael Elliott. 

### Filtering code (Version 4)

```{r}
library(dplyr)

filtering <- function(df) {
  
  # Remove unnecessary column
  df <- df %>% select(-Systolic, -Diastolic, -MAP, -HR, -`BV*`, -CO, -LVET, -SV, -`AS*`)
  
  # Convert IBI to seconds
  df <- df %>% mutate(IBI = IBI / 1000)
  print("Converted IBI to seconds:")
  
  # Remove if the first row IBI is zero
  df <- if (df$IBI[1] == 0) df[-1, ] else df

  # Helper function to compute E_10
  compute_E_10 <- function(df, i) {
    if (i + 10 > nrow(df) || i - 1 < 1) return(NA)
    sum(abs((df$IBI[i:(i + 10)] - df$IBI[(i - 1):(i + 9)]) / df$IBI[(i - 1):(i + 9)])) / 10
  }
  
  # Process the data
  i <- 1
  while (i < nrow(df)) {
    print(paste("Processing row:", i))
    
    # If IBI > 1.3 s, then remove IBI_i and its timestamp
    if (!is.na(df$IBI[i]) && df$IBI[i] > 1.3) {
      print(paste("IBI >", 1.3, "at row", i, "- Removing row"))
      df <- df[-i, ]
      next
    }
    
    # if IBI < 0.3s, then do all the merging process
    if (!is.na(df$IBI[i]) && df$IBI[i] < 0.3) {
      print(paste("IBI <", 0.3, "at row", i, "- Performing merge process"))
      
      if (i + 1 > nrow(df)) break
      IBI_r <- df$IBI[i] + df$IBI[i + 1]
      ER_r <- if (!is.na(df$IBI[i]) && !is.na(IBI_r)) abs((IBI_r - df$IBI[i]) / df$IBI[i]) else NA
      ER_l <- if (i > 1 && !is.na(df$IBI[i - 1]) && !is.na(IBI_r)) abs((IBI_r - df$IBI[i - 1]) / df$IBI[i - 1]) else NA
      E_tot_r <- if (!is.na(ER_r) && !is.na(ER_l)) ER_r + ER_l else NA
      E_10 <- compute_E_10(df, i)
      
      print(paste("IBI_r:", IBI_r, "ER_r:", ER_r, "ER_l:", ER_l, "E_tot_r:", E_tot_r, "E_10:", E_10))
      
      # If IBI_r < 1.3s and ER_l <= E_10 and ER_r <= E_10 then 
      # do right merge, replace IBI_i+1 by IBI_r and delete IBI_i and its timestamp
      if (!is.na(IBI_r) && !is.na(ER_r) && !is.na(ER_l) && !is.na(E_10) && IBI_r < 1.3 && ER_l <= E_10 && ER_r <= E_10) {
        df$IBI[i + 1] <- IBI_r
        df <- df[-i, ]
        print("Performed right merge:")
        next
      } else {
        if (i - 1 < 1) break
        IBI_l <- df$IBI[i] + df$IBI[i - 1]
        EL_r <- if (!is.na(df$IBI[i]) && !is.na(IBI_l)) abs((IBI_l - df$IBI[i]) / df$IBI[i]) else NA
        EL_l <- if (!is.na(df$IBI[i - 1]) && !is.na(IBI_l)) abs((IBI_l - df$IBI[i - 1]) / df$IBI[i - 1]) else NA
        E_tot_l <- if (!is.na(EL_r) && !is.na(EL_l)) EL_r + EL_l else NA
        E_10_prev <- compute_E_10(df, i - 1)
        
        print(paste("IBI_l:", IBI_l, "EL_r:", EL_r, "EL_l:", EL_l, "E_tot_l:", E_tot_l, "E_10_prev:", E_10_prev))
        
        # If IBI_l < 1.3 s and EL_l <= E10 and EL_r <= E_10 then 
        # left merge, replace IBI_i by IBI_l and delete IBI_i-1 and its timestamp
        if (!is.na(IBI_l) && !is.na(EL_l) && !is.na(EL_r) && !is.na(E_10_prev) && IBI_l < 1.3 && EL_l <= E_10_prev && EL_r <= E_10_prev) {
          df$IBI[i - 1] <- IBI_l
          df <- df[-i, ]
          print("Performed left merge:")
          next
        } else if (!is.na(IBI_r) && !is.na(IBI_l) && IBI_r > 1.3 && IBI_l > 1.3) {
          if (i + 1 <= nrow(df)) { # Else if IBI_r > 1.3 and IBI_l > 1.3, then Delete both {IBI}_i and {IBI}_{i+1}
            df <- df[-c(i, i + 1), ]
          } else {
            df <- df[-i, ]
          }
          print("Deleted rows due to IBI_r > 1.3 and IBI_l > 1.3:")
          next
        } 
        # Else if IBI_r < 1.3 and IBI_l > 1.3 then Replace {IBI}_{i+1} by {IBI}_r and delete {IBI}_i and its timestamp
        else if (!is.na(IBI_r) && !is.na(IBI_l) && IBI_r < 1.3 && IBI_l > 1.3) {
          df$IBI[i + 1] <- IBI_r 
          df <- df[-i, ]
          print("Replaced IBI_i+1 by IBI_r and deleted IBI_i:")
          next
        } 
        # Else if IBI_r > 1.3s and IBI_l < 1.3s then Replace {IBI}_{i-1} by {IBI}_l and delete {IBI}_i and its timestamp
        else if (!is.na(IBI_r) && !is.na(IBI_l) && IBI_r > 1.3 && IBI_l < 1.3) {
          df$IBI[i - 1] <- IBI_l 
          df <- df[-i, ]
          print("Replaced IBI_i-1 by IBI_l and deleted IBI_i:")
          next
        } 
        # Else if both {IBI}_l and {IBI}_r < 1.3s but errors {Etot}_r and {Etot}_l are higher than 0.4, then keep the one with smaller error
        else if (!is.na(IBI_l) && !is.na(IBI_r) && !is.na(E_tot_r) && !is.na(E_tot_l) && IBI_l < 1.3 && IBI_r < 1.3 && E_tot_r > 0.4 
                 && E_tot_l > 0.4) {
          if (E_tot_r < E_tot_l) { 
            df$IBI[i + 1] <- IBI_r
            df <- df[-i, ]
          } else {
            df$IBI[i - 1] <- IBI_l
            df <- df[-i, ]
          }
          print("Kept the one with smaller error and removed the other:")
          next
        }
      }
    }
    # restate the iteration
    i <- i + 1
  }
  
  # Handle the last row
  # if the IBI < 0.3
  while (nrow(df) > 0 && df$IBI[nrow(df)] < 0.3) {
    
    # compute IBI left
    IBI_l <- df$IBI[nrow(df)] + df$IBI[nrow(df) - 1]
    
    # if IBI_l > 1.3 s
    if (IBI_l > 1.3) {
      df <- df[-nrow(df), ] # remove the row
      print("Removed last row due to IBI < 0.3 and IBI_l > 1.3:")
      next
    } else {
      df$IBI[nrow(df) - 1] <- IBI_l # otherwise, do left merge
      df <- df[-nrow(df), ]
      print("Performed left merge on the last row:")
    }
  }
  
  # if IBI > 1.3
  if (nrow(df) > 0 && df$IBI[nrow(df)] > 1.3) {
    df <- df[-nrow(df), ] # remove the row
    print("Removed last row due to IBI > 1.3:")
  }
  
  # Convert back RR to milliseconds
  df <- df %>% mutate(IBI = IBI * 1000)
  print("Converted RR back to milliseconds:")
  
  return(df)
}
```

### Impute IBI version 2 (after debugging) 

When I apply the previous code, the loop seems not converge. Therefore, I revise the code with the details below:

- Compute E_10 Once: The E_10 value is computed once for each gap instead of recalculating during each retry.

- Retry Strategy: The code retries generating the new IBI value up to max_attempts times, with a slight loosening of the standard deviation (sigma) each time.

- Skip Condition: If the maximum number of retries is reached, the function skips to the next interval, preventing the loop from being stuck indefinitely.

```{r}
impute_IBI_intervals <- function(df) {
  
  # library
  library(dplyr)
  library(tidyr)
  
  # Separate time into hours, minutes, seconds, and milliseconds, then calculate Timestamps
  df <- df %>%
    mutate(Temp = Time) %>%  # Create a temporary column to split
    separate(Temp, into = c("hours", "minutes", "sec_millisec"), sep = ":") %>%
    separate(sec_millisec, into = c("seconds", "milliseconds"), sep = "\\.") %>%
    mutate(across(c(hours, minutes, seconds, milliseconds), as.integer)) %>%
    mutate(
      hours = hours * 60 * 60 * 1000,
      minutes = minutes * 60 * 1000,
      seconds = seconds * 1000,
      Timestamps = hours + minutes + seconds + milliseconds
  ) %>%
  select(-hours, -minutes, -seconds, -milliseconds)  # Remove intermediate columns
  
  # Convert IBI and Time from milliseconds to seconds
  df <- df %>% mutate(IBI = IBI / 1000, Timestamps = Timestamps / 1000)
  
  # Helper function to compute E_10
  compute_E_10 <- function(df, i) {
    if (i + 10 > nrow(df) || i - 1 < 1) return(NA)
    sum(abs((df$IBI[i:(i + 10)] - df$IBI[(i - 1):(i + 9)]) / df$IBI[(i - 1):(i + 9)])) / 10
  }
  
  # Helper function to compute deviation E_r and E_l
  compute_deviation <- function(IBI_j, IBI_prev, IBI_next) {
    E_l <- abs((IBI_j - IBI_prev) / IBI_prev)
    E_r <- abs((IBI_next - IBI_j) / IBI_j)
    return(list(E_l = E_l, E_r = E_r))
  }
  
  # Iterate over IBI intervals
  i <- 1
  while (i < nrow(df)) {
    cat("Iteration:", i, "\n")
    
    # Check if the time difference > 1.3 and not equal to IBI[i+1]
    while (!is.na(df$Timestamps[i]) && !is.na(df$Timestamps[i + 1]) && (df$Timestamps[i + 1] - df$Timestamps[i]) > 1.3) {
      
      # Skip the gap if there are not enough previous IBI intervals
      if (i - 9 <= 0) {
        cat("Not enough previous IBI intervals, skipping gap.\n")
        break
      }
      
      # Calculate E_10 once per gap
      E_10 <- compute_E_10(df, i)
      
      attempt_count <- 0
      max_attempts <- 5  # Set a reasonable limit
      while (attempt_count < max_attempts) {
        # Use mean and std deviation from 10 previous observations
        mu <- mean(df$IBI[(i - 9):i])
        sigma <- sd(df$IBI[(i - 9):i])
        
        # Generate new IBI value
        # set.seed(999)
        new_IBI <- rnorm(1, mean = mu, sd = sigma)
        cat("Generated new IBI:", new_IBI, "\n")
        
        # Insert the new IBI value to the gap
        df <- df %>% add_row(IBI = new_IBI, .before = i + 1)
        
        # Insert new Timestamps to the inserted new IBI
        T_end <- df$Timestamps[i + 2]
        new_time <- T_end - df$IBI[i + 2]
        df$Timestamps[i + 1] <- new_time
        cat("New Timestamps value inserted:", new_time, "\n")
        
        # Check conditions for the new IBI
        deviations <- compute_deviation(new_IBI, df$IBI[i - 1], df$IBI[i + 1])
        
        if (!is.na(deviations$E_l) && !is.na(deviations$E_r) && !is.na(E_10) &&
            deviations$E_l <= E_10 && deviations$E_r <= E_10 &&
            new_IBI >= 0.3 && new_IBI <= 1.3) {
          # Conditions are met, continue to next gap
          break
        }
        
        # Conditions not met, remove the inserted IBI and try again
        cat("Conditions not met, retrying...\n")
        df <- df[-(i + 1), ] # Remove the observation
        attempt_count <- attempt_count + 1
        sigma <- sigma * 1.05  # Loosen the requirements slightly
      }
      
      # If maximum attempts reached without success, skip to next interval
      if (attempt_count == max_attempts) {
        cat("Exceeded maximum attempts. Skipping this gap.\n")
        break
      }
    }
    i <- i + 1
  }
  
  # Convert IBI and Time from seconds back to milliseconds
  df <- df %>% mutate(IBI = round((IBI * 1000),0), Timestamps = round((Timestamps * 1000),0))
  
  # Add column about imputation flag, 1 is imputed, 0 is not imputed.
  df$Flag <- ifelse(is.na(df$Time), 1, 0)
  
  # Fill out the empty Time 
  df <- df %>%
  mutate(
    Time = ifelse(is.na(Time), # check NA value
                  sprintf("%02d:%02d:%02d.%03d", 
                          Timestamps %/% (60*60*1000),                      # Hours
                          (Timestamps %/% (60*1000)) %% 60,                  # Minutes
                          (Timestamps %/% 1000) %% 60,                       # Seconds
                          Timestamps %% 1000),                               # Milliseconds
                  Time)  # If not NA, keep the existing value
  )
  
  # Fill out the Date
  df <- df %>%
  fill(Date, .direction = "downup")  # Fills down first, then up if needed
  
  # Return the dataframe result
  return(df)
}
```

### Test dataset

```{r}
# input the directory
directory <- "O:/WorkLife-Data/biomonitors/caretaker_cbp/59503/Type_2_Data" 
# change the directory with the current directory

# input the file path
file <- c("bbb_vitals_2024-06-29_13-02-26.csv") 
# change the directory with the current directory

# file path
full_file_path <- file.path(directory, file)
  
# read the CSV file
data_original59503 <- data.table::fread(full_file_path)
```

```{r}
# apply the helper function to one dataset
data_filtered59503 <- filtering(data_original59503) 
```

```{r}
# subset the data (turned off for testing purpose)
# check for smaller dataset such as 300, 1000, 5000, 15000, ...., max observation
# extract4a <- data_filtered59503[1:300, ]
# imputed_extract4a <- impute_IBI_intervals(extract4a)
```

```{r}
# apply the imputation code
data_imputed59503 <- impute_IBI_intervals(data_filtered59503)
```

```{r}
# check the summary of filtered data
summary(data_original59503$IBI)
dim(data_original59503)
head(data_original59503)

# check the summary of data before imputation
summary(data_filtered59503$IBI)
dim(data_filtered59503)
head(data_filtered59503)

# check the summary of data after imputation
summary(data_imputed59503$Timestamps)
summary(data_imputed59503$IBI)
dim(data_imputed59503)
head(data_imputed59503)
```

```{r}
# save the imputed datasets
output_file <- paste0("59503_imputed.txt")

# Full output file path
output_file_path <- file.path(directory, output_file)

# Save the processed data
write.table(data_imputed59503, file = output_file_path, row.names = FALSE, col.names = TRUE)
```

### Apply the imputation function to the dataset

Note: You need to change the file path with the current file path of the data you are currently working on. 

### Participant 38190

```{r}
# input the directory
directory <- "O:/WorkLife-Data/biomonitors/caretaker_cbp/38190/Type_2_Data" 
# change the directory with the current directory

# input the file path
file <- c("bbb_vitals_2024-07-06_13-10-54.csv") 
# change the directory with the current directory

# file path
full_file_path <- file.path(directory, file)
  
# read the CSV file
data_original38190 <- data.table::fread(full_file_path)
```

```{r}
# apply the helper function to one dataset
data_filtered38190 <- filtering(data_original38190) 
```

```{r}
# subset the data (turned off for testing purpose)
# check for smaller dataset such as 300, 1000, 5000, 15000, ...., max observation
# extract4a <- data_filtered[1:190674, ]
```

```{r}
# apply the imputation code
data_imputed38190 <- impute_IBI_intervals(data_filtered38190)
```

```{r}
# check the summary of filtered data
summary(data_original38190$IBI)
dim(data_original38190)
head(data_original38190)

# check the summary of data before imputation
summary(data_filtered38190$IBI)
dim(data_filtered38190)
head(data_filtered38190)

# check the summary of data after imputation
summary(data_imputed38190$Timestamps)
summary(data_imputed38190$IBI)
dim(data_imputed38190)
head(data_imputed38190)
```

```{r}
# save the imputed datasets
output_file <- paste0("38190_imputed.txt")

# Full output file path
output_file_path <- file.path(directory, output_file)

# Save the processed data
write.table(data_imputed38190, file = output_file_path, row.names = FALSE, col.names = TRUE)
```

### Participant 19686

```{r}
# input the directory
directory <- "O:/WorkLife-Data/biomonitors/caretaker_cbp/19686/Type_2_Data" 
# change the directory with the current directory

# input the file path
file <- c("bbb_vitals_2024-07-17_13-44-34.csv") 
# change the directory with the current directory

# file path
full_file_path <- file.path(directory, file)
  
# read the CSV file
data_original19686 <- data.table::fread(full_file_path)
```

```{r}
# apply the helper function to one dataset
data_filtered19686 <- filtering(data_original19686) 
```

```{r}
# subset the data (turned off for testing purpose)
# check for smaller dataset such as 300, 1000, 5000, 15000, ...., max observation
# extract4a <- data_filtered[1:190674, ]
```

```{r}
# apply the imputation code
data_imputed19686 <- impute_IBI_intervals(data_filtered19686)
```

```{r}
# check the summary of filtered data
summary(data_original19686$IBI)
dim(data_original19686)
head(data_original19686)

# check the summary of data before imputation
summary(data_filtered19686$IBI)
dim(data_filtered19686)
head(data_filtered19686)

# check the summary of data after imputation
summary(data_imputed19686$Timestamps)
summary(data_imputed19686$IBI)
dim(data_imputed19686)
head(data_imputed19686)
```

```{r}
# save the imputed datasets
output_file <- paste0("19686_imputed.txt")

# Full output file path
output_file_path <- file.path(directory, output_file)

# Save the processed data
write.table(data_imputed19686, file = output_file_path, row.names = FALSE, col.names = TRUE)
```

### Code Archive

```{r}
impute_IBI_intervals <- function(df) {
  
  # library
  library(dplyr)
  library(tidyr)
  
  # Separate time into hours, minutes, seconds, and milliseconds, then calculate Timestamps
  df <- df %>%
    mutate(Temp = Time) %>%  # Create a temporary column to split
    separate(Temp, into = c("hours", "minutes", "sec_millisec"), sep = ":") %>%
    separate(sec_millisec, into = c("seconds", "milliseconds"), sep = "\\.") %>%
    mutate(across(c(hours, minutes, seconds, milliseconds), as.integer)) %>%
    mutate(
      hours = hours * 60 * 60 * 1000,
      minutes = minutes * 60 * 1000,
      seconds = seconds * 1000,
      Timestamps = hours + minutes + seconds + milliseconds
  ) %>%
  select(-hours, -minutes, -seconds, -milliseconds)  # Remove intermediate columns
  
  # Convert IBI and Time from milliseconds to seconds
  df <- df %>% mutate(IBI = IBI / 1000, Timestamps = Timestamps / 1000)
  
  # Helper function to compute E_10
  compute_E_10 <- function(df, i) {
    if (i + 10 > nrow(df) || i - 1 < 1) return(NA)
    sum(abs((df$IBI[i:(i + 10)] - df$IBI[(i - 1):(i + 9)]) / df$IBI[(i - 1):(i + 9)])) / 10
  }
  
  # Helper function to compute deviation E_r and E_l
  compute_deviation <- function(IBI_j, IBI_prev, IBI_next) {
    E_l <- abs((IBI_j - IBI_prev) / IBI_prev)
    E_r <- abs((IBI_next - IBI_j) / IBI_j)
    return(list(E_l = E_l, E_r = E_r))
  }
  
  # Iterate over IBI intervals
  i <- 1
  while (i < nrow(df)) {
    cat("Iteration:", i, "\n")
    
    # Check if the time difference > 1.3 and not equal to IBI[i+1]
    while (!is.na(df$Timestamps[i]) && !is.na(df$Timestamps[i + 1]) && (df$Timestamps[i + 1] - df$Timestamps[i]) > 1.3) {
      
      # Skip the gap if there are not enough previous IBI intervals
      if (i - 9 <= 0) {
        cat("Not enough previous IBI intervals, skipping gap.\n")
        break
      }
      
      attempt_count <- 0
      inserted_data_count <- 0
      repeat {
        # Use mean and std deviation from 10 previous observations
        mu <- mean(df$IBI[(i - 9):i])
        sigma <- sd(df$IBI[(i - 9):i])
        
        # Generate new IBI value
        new_IBI <- rnorm(1, mean = mu, sd = sigma)
        cat("Generated new IBI:", new_IBI, "\n")
        
        # Insert the new IBI value to the gap
        df <- df %>% add_row(IBI = new_IBI, .before = i + 1)
        inserted_data_count <- inserted_data_count + 1
        
        # Insert new Timestamps to the inserted new IBI
        T_end <- df$Timestamps[i + 2]
        new_time <- T_end - df$IBI[i + 2]
        df$Timestamps[i + 1] <- new_time
        cat("New Timestamps value inserted:", new_time, "\n")
        
        # Check conditions for the new IBI
        deviations <- compute_deviation(new_IBI, df$IBI[i - 1], df$IBI[i + 1])
        E_10 <- compute_E_10(df, i)
        
        if (!is.na(deviations$E_l) && !is.na(deviations$E_r) && !is.na(E_10) && deviations$E_l <= E_10 &&  deviations$E_r <= E_10 &&
            new_IBI >= 0.3 && new_IBI <= 1.3) {
          # Conditions are met, continue to next gap
          break
        }
        
        # Conditions not met, retry
        attempt_count <- attempt_count + 1
        if (attempt_count > 3) {
          cat("Exceeded maximum attempts.\n")
          if (inserted_data_count == 1) {
            df <- df[-(i + 1), ] # Remove one observation
          } else {
            df <- df[-((i + 1):(i + 2)), ] # Remove two observations
            for (j in 1:4) {
              E_10 <- E_10 * 1.05 # Loosen the requirements by 5%
            }
          }
          break
        }
        
        cat("Conditions not met, retrying...\n")
        df <- df[-(i + 1), ] # Remove observation
        inserted_data_count <- inserted_data_count - 1
        E_10 <- E_10 * 1.05 # Loosen the requirements by 5%
      }
    }
    i <- i + 1
  }
  
  # Convert IBI and Time from seconds back to milliseconds
  df <- df %>% mutate(IBI = IBI * 1000, Timestamps = Timestamps * 1000)
  
  # Return the dataframe result
  return(df)
}

```
