---
title: "Filtering_Imputation"
author: "Aulia Dini Rafsanjani"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This file contains final code of two parts: Filtering and Imputation. The filtering process is handled by the `filtering` function, while the imputation process is managed by the `impute_rr_intervals` function. To run the code, first execute the `filtering` code chunks and apply the filtering function to your datasets. After filtering, you can use the resulting data_filtered as input for the `impute_rr_intervals` steps.

Please note that the processing time may take a few minutes to complete. Therefore, I recommend running this code for each dataset individually (rather than many datasets simultaneously) to avoid potential errors.

The file naming in this code still uses the Dropbox naming format. To apply this code to real participants, you should use the new naming convention format in the O drive. 

### Filtering code

### Filtering 4 

I add the conditions above to filter the last row of data.

```{r}
library(dplyr)

filtering <- function(df) {
  
  # Convert RR to seconds
  df <- df %>% mutate(RR = RR / 1000)
  
  # Helper function to compute E_10
  compute_E_10 <- function(df, i) {
    if (i + 10 > nrow(df) || i - 1 < 1) return(NA)
    sum(abs((df$RR[i:(i + 10)] - df$RR[(i - 1):(i + 9)]) / df$RR[(i - 1):(i + 9)])) / 10
  }
  
  # Process the data
  # set the starting point
  i <- 1
  while (i < nrow(df)) {
    
    # If RR > 1.3 s, then remove RR_i and its timestamp
    # this is the new condition
    if (!is.na(df$RR[i]) && df$RR[i] > 1.3) {
      df <- df[-i, ]
      next
    }
    
    # if RR < 0.3s, then do all the merging process
    if (!is.na(df$RR[i]) && df$RR[i] < 0.3) {
      if (i + 1 > nrow(df)) break
      RR_r <- df$RR[i] + df$RR[i + 1]
      ER_r <- abs((RR_r - df$RR[i]) / df$RR[i])
      ER_l <- abs((RR_r - df$RR[i - 1]) / df$RR[i - 1])
      E_tot_r <- if (!is.na(ER_r) && !is.na(ER_l)) ER_r + ER_l else NA
      E_10 <- compute_E_10(df, i)
      
      # If RR_r < 1.3s and ER_l <= E_10 and ER_r <= E_10 then 
      # do right merge, replace RR_i+1 by RR_r and delete RR_i and its timestamp
      if (!is.na(RR_r) && !is.na(ER_r) && !is.na(ER_l) && !is.na(E_10) && RR_r < 1.3 && ER_l <= E_10 && ER_r <= E_10) {
        df$RR[i + 1] <- RR_r
        df <- df[-i, ]
        next
      } else {
        if (i - 1 < 1) break
        RR_l <- df$RR[i] + df$RR[i - 1]
        EL_r <- abs((RR_l - df$RR[i]) / df$RR[i])
        EL_l <- abs((RR_l - df$RR[i - 1]) / df$RR[i - 1])
        E_tot_l <- if (!is.na(EL_r) && !is.na(EL_l)) EL_r + EL_l else NA
        E_10_prev <- compute_E_10(df, i - 1)
        
        # If RR_l < 1.3 s and EL_l <= E10 and EL_r <= E_10 then 
        # left merge, replace RR_i by RR_l and delete RR_i-1 and its timestamp
        if (!is.na(RR_l) && !is.na(EL_l) && !is.na(EL_r) && !is.na(E_10_prev) && RR_l < 1.3 && EL_l <= E_10_prev && EL_r <= E_10_prev) {
          df$RR[i - 1] <- RR_l
          df <- df[-i, ]
          next
        } else if (!is.na(RR_r) && !is.na(RR_l) && RR_r > 1.3 && RR_l > 1.3) {
          if (i + 1 <= nrow(df)) { # Else if RR_r > 1.3 and RR_l > 1.3, then Delete both {RR}_i and {RR}_{i+1}
            df <- df[-c(i, i + 1), ]
          } else {
            df <- df[-i, ]
          }
          next
        } 
        # Else if RR_r < 1.3 and RR_l > 1.3 then Replace {RR}_{i+1} by {RR}_r and delete {RR}_i and its timestamp
        else if (!is.na(RR_r) && !is.na(RR_l) && RR_r < 1.3 && RR_l > 1.3) {
          df$RR[i + 1] <- RR_r 
          df <- df[-i, ]
          next
        } 
        # Else if RR_r > 1.3s and RR_l < 1.3s then Replace {RR}_{i-1} by {RR}_l and delete {RR}_i and its timestamp
        else if (!is.na(RR_r) && !is.na(RR_l) && RR_r > 1.3 && RR_l < 1.3) {
          df$RR[i - 1] <- RR_l 
          df <- df[-i, ]
          next
        } 
        # Else if both {RR}_l and {RR}_r<\ 1.3\ s but errors {Etot}_r and {Etot}_l are higher than 0.4, then keep the one with smaller error
        else if (!is.na(RR_l) && !is.na(RR_r) && !is.na(E_tot_r) && !is.na(E_tot_l) && RR_l < 1.3 && RR_r < 1.3 && E_tot_r > 0.4 
                 && E_tot_l > 0.4) {
          if (E_tot_r < E_tot_l) { 
            df$RR[i + 1] <- RR_r
            df <- df[-i, ]
          } else {
            df$RR[i - 1] <- RR_l
            df <- df[-i, ]
          }
          next
        }
      }
    }
    # restate the iteration
    i <- i + 1
  }
  
  # Handle the last row
  # if the RR < 0.3
  while (nrow(df) > 0 && df$RR[nrow(df)] < 0.3) {
    
    # compute RR left
    RR_l <- df$RR[nrow(df)] + df$RR[nrow(df)-1]
    
    # if RR_l > 1.3 s
    if (RR_l > 1.3) {
      df <- df[-nrow(df), ] # remove the row
      next
    } else {
      df$RR[nrow(df)-1] <- RR_l # otherwise, do left merge
      df <- df[-nrow(df), ]
    }
  }
  
  # if RR > 1.3
  if (nrow(df) > 0 && df$RR[nrow(df)] > 1.3) {
    df <- df[-nrow(df), ] # remove the row
  }
  
  # Convert back RR to milliseconds
  df <- df %>% mutate(RR = RR * 1000)
  
  # Return the dataframe result
  return(df)
}
```

### Imputation code

### Final version

```{r}
library(dplyr)

impute_rr_intervals <- function(df) {
  
  # Convert RR and Time from milliseconds to seconds
  df <- df %>% mutate(RR = RR / 1000, Time = Time / 1000)
  
  # Helper function to compute E_10
  compute_E_10 <- function(df, i) {
    if (i + 10 > nrow(df) || i - 1 < 1) return(NA)
    sum(abs((df$RR[i:(i + 10)] - df$RR[(i - 1):(i + 9)]) / df$RR[(i - 1):(i + 9)])) / 10
  }
  
  # Helper function to compute deviation E_r and E_l
  compute_deviation <- function(RR_j, RR_prev, RR_next) {
    E_l <- abs((RR_j - RR_prev) / RR_prev)
    E_r <- abs((RR_next - RR_j) / RR_j)
    return(list(E_l = E_l, E_r = E_r))
  }
  
  # Iterate over RR intervals
  i <- 1
  while (i < nrow(df)) {
    cat("Iteration:", i, "\n")
    
    # Check if the time difference > 1.3 and not equal to RR[i+1]
    while (!is.na(df$Time[i]) && !is.na(df$Time[i + 1]) && (df$Time[i + 1] - df$Time[i]) > 1.3) {
      
      # Skip the gap if there are not enough previous RR intervals
      if (i - 9 <= 0) {
        cat("Not enough previous RR intervals, skipping gap.\n")
        break
      }
      
      attempt_count <- 0
      repeat {
        # Use mean and std deviation from 10 previous observations
        mu <- mean(df$RR[(i - 9):i])
        sigma <- sd(df$RR[(i - 9):i])
        
        # Generate new RR value
        new_rr <- rnorm(1, mean = mu, sd = sigma)
        cat("Generated new RR:", new_rr, "\n")
        
        # Insert the new RR value to the gap
        df <- df %>% add_row(RR = new_rr, .before = i + 1)
        
        # Insert new Time to the inserted new RR
        T_end <- df$Time[i + 2]
        new_time <- T_end - df$RR[i + 2]
        df$Time[i + 1] <- new_time
        cat("New Time value inserted:", new_time, "\n")
        
        # Check conditions for the new RR
        deviations <- compute_deviation(new_rr, df$RR[i - 1], df$RR[i + 1])
        E_10 <- compute_E_10(df, i)
        
        if (new_rr > 0.3 && new_rr < 1.3) {
          cat("Conditions met for new RR.\n")
          break
        } else {
          cat("Conditions not met, retrying...\n")
          df <- df[-(i + 1), ] # Remove observation
          E_10 <- E_10 * 1.05 # Loosen the requirements
          attempt_count <- attempt_count + 1
          if (attempt_count > 3) {
            cat("Exceeded maximum attempts.\n")
            break
          }
        }
      }
    }
    i <- i + 1
  }
  
  # Convert RR and Time from seconds back to milliseconds
  df <- df %>% mutate(RR = RR * 1000, Time = Time * 1000)
  
  # Return the dataframe result
  return(df)
}

```

### Test the imputation function to the dataset

### Test data to other datasets

```{r}
directory <- "/Users/auliadinirafsanjani/Dropbox (University of Michigan)/WorkLife_ECGTest/Test"

file <- c("test867601_ADVHRV_2024-05-07.txt")

#  file path
full_file_path <- file.path(directory, file)
  
# Read the CSV file
data_original <- data.table::fread(full_file_path, sep = "\t")
  
# change column name
colnames(data_original)[4] <- "conv_type"
```

```{r}
# apply the helper function to one dataset
data_filtered<- filtering(data_original) 
```

```{r}
# subset the data
extract4a <- data_filtered[1:444631, ]
```

```{r}
# apply the imputation code
data_imputed_4b <- impute_rr_intervals(extract4a)
```

```{r}
# check the summary of data before imputation
summary(extract4a$Time)
summary(extract4a$RR)
dim(extract4a)
head(extract4a)

# check the summary of data after imputation
summary(data_imputed_4b$Time)
summary(data_imputed_4b$RR)
dim(data_imputed_4b)
head(data_imputed_4b)
```
Findings:

0. Test data : Ben, Neil, MA, Dini, Amanda, Jenny, Mary, Angelica

1. The code works well on the small and large datasets for all tests participant: Angelica, Ben, Neil, Marie-Anne, Karis, Jenny, Amanda, Aulia. There is not any observation that is not within range of 0.3s < RR < 1.3s. 

2. The mean and standard deviation from the ten first observation issue --> Skip and do not apply imputation if the prior observations are less than 10. 

3. Checking conditions do not converge for conditions 2 and 3 --> I remove the physiological conditions 2 and 3 (it is justified by the paper page 6). Therefore, the conditions checking for imputation only check whether 0.3s < RR <1.3s. 