---
title: "ECG_Filtering_Imputation"
author: "Aulia Dini Rafsanjani"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This file contains final code of two parts: Filtering and Imputation for ECG data. The filtering process is handled by the `filtering` function, while the imputation process is managed by the `impute_RR_intervals` function. To run the code, first execute the `filtering` code chunks and apply the filtering function to your datasets. After filtering, you can use the resulting data_filtered as input for the `impute_rr_intervals` steps.

Please note that the processing time may take a few minutes to complete. Therefore, I recommend running this code for each dataset individually (rather than many datasets simultaneously) to avoid potential errors.

The file name / file path in this code still uses the Dropbox naming format. To apply this code to real participants, you should use the new naming convention format in the O drive. 

Side Notes :

- The imputation code started from the eleventh row. In this case, if there is a time gap from the first to the tenth observation, the code will automatically skip them. This is because the algorithm uses the variance from the previous ten observations. Therefore, the best decision (so far) is to skip the first to tenth observations. 

- The imputation code contains set.seed () function, and I am currently commenting on it. The decision to include the code of set.seed() will follow the discussion results with the statistics expert (Michael Elliott).  

### Filtering code (Version 4)

```{r}
library(dplyr)

filtering <- function(df) {
  
  # Convert RR to seconds
  df <- df %>% mutate(RR = RR / 1000)
  print("Converted RR to seconds:")
  print(head(df))
  
  # Helper function to compute E_10
  compute_E_10 <- function(df, i) {
    if (i + 10 > nrow(df) || i - 1 < 1) return(NA)
    sum(abs((df$RR[i:(i + 10)] - df$RR[(i - 1):(i + 9)]) / df$RR[(i - 1):(i + 9)])) / 10
  }
  
  # Process the data
  i <- 1
  while (i <= nrow(df)) {
    print(paste("Iteration:", i))
    
    # If RR > 1.3 s, then remove RR_i and its timestamp
    if (!is.na(df$RR[i]) && df$RR[i] > 1.3) {
      print(paste("Removing row with RR >", 1.3, "at index", i))
      df <- df[-i, ]
      next
    }
    
    # if RR < 0.3s, then do all the merging process
    if (!is.na(df$RR[i]) && df$RR[i] < 0.3) {
      print(paste("Processing RR <", 0.3, "at index", i))
      if (i + 1 > nrow(df)) {
        print(paste("No RR_r available for RR <", 0.3, "at index", i))
        i <- i + 1
        next
      }
      
      RR_r <- df$RR[i] + df$RR[i + 1]
      ER_r <- abs((RR_r - df$RR[i]) / df$RR[i])
      ER_l <- if (i > 1) abs((RR_r - df$RR[i - 1]) / df$RR[i - 1]) else NA
      E_tot_r <- if (!is.na(ER_r) && !is.na(ER_l)) ER_r + ER_l else NA
      E_10 <- compute_E_10(df, i)
      print(paste("RR_r:", RR_r, "ER_r:", ER_r, "ER_l:", ER_l, "E_tot_r:", E_tot_r, "E_10:", E_10))
      
      # If RR_r < 1.3s and ER_l <= E_10 and ER_r <= E_10 then do right merge
      if (!is.na(RR_r) && !is.na(ER_r) && !is.na(ER_l) && !is.na(E_10) && RR_r < 1.3 && ER_l <= E_10 && ER_r <= E_10) {
        print(paste("Right merging at index", i))
        df$RR[i + 1] <- RR_r
        df <- df[-i, ]
        next
      } else {
        if (i == 1) {
          print("Skipping left merge process at the first index")
          i <- i + 1
          next
        }
        RR_l <- df$RR[i] + df$RR[i - 1]
        EL_r <- abs((RR_l - df$RR[i]) / df$RR[i])
        EL_l <- abs((RR_l - df$RR[i - 1]) / df$RR[i - 1])
        E_tot_l <- if (!is.na(EL_r) && !is.na(EL_l)) EL_r + EL_l else NA
        E_10_prev <- compute_E_10(df, i - 1)
        print(paste("RR_l:", RR_l, "EL_r:", EL_r, "EL_l:", EL_l, "E_tot_l:", E_tot_l, "E_10_prev:", E_10_prev))
        
        # If RR_l < 1.3 s and EL_l <= E_10 and EL_r <= E_10 then left merge
        if (!is.na(RR_l) && !is.na(EL_l) && !is.na(EL_r) && !is.na(E_10_prev) && RR_l < 1.3 && EL_l <= E_10_prev && EL_r <= E_10_prev) {
          print(paste("Left merging at index", i))
          df$RR[i - 1] <- RR_l
          df <- df[-i, ]
          next
        } else if (!is.na(RR_r) && !is.na(RR_l) && RR_r > 1.3 && RR_l > 1.3) {
          if (i + 1 <= nrow(df)) { 
            print(paste("Deleting both RR_i and RR_{i+1} at index", i))
            df <- df[-c(i, i + 1), ]
          } else {
            print(paste("Deleting RR_i at index", i))
            df <- df[-i, ]
          }
          next
        } else if (!is.na(RR_r) && !is.na(RR_l) && RR_r < 1.3 && RR_l > 1.3) {
          print(paste("Replacing RR_{i+1} by RR_r and deleting RR_i at index", i))
          df$RR[i + 1] <- RR_r 
          df <- df[-i, ]
          next
        } else if (!is.na(RR_r) && !is.na(RR_l) && RR_r > 1.3 && RR_l < 1.3) {
          print(paste("Replacing RR_{i-1} by RR_l and deleting RR_i at index", i))
          df$RR[i - 1] <- RR_l 
          df <- df[-i, ]
          next
        } else if (!is.na(RR_l) && !is.na(RR_r) && !is.na(E_tot_r) && !is.na(E_tot_l) && RR_l < 1.3 && RR_r < 1.3 && E_tot_r > 0.4 && E_tot_l > 0.4) {
          if (E_tot_r < E_tot_l) { 
            print(paste("Keeping RR_r with smaller error at index", i))
            df$RR[i + 1] <- RR_r
            df <- df[-i, ]
          } else {
            print(paste("Keeping RR_l with smaller error at index", i))
            df$RR[i - 1] <- RR_l
            df <- df[-i, ]
          }
          next
        }
      }
    }
    i <- i + 1
  }
  
  # Handle the last row
  while (nrow(df) > 0 && df$RR[nrow(df)] < 0.3) {
    print("Handling last row with RR <", 0.3)
    RR_l <- df$RR[nrow(df)] + df$RR[nrow(df)-1]
    if (RR_l > 1.3) {
      print("Removing last row with RR >", 1.3)
      df <- df[-nrow(df), ]
    } else {
      print("Left merging last row")
      df$RR[nrow(df)-1] <- RR_l
      df <- df[-nrow(df), ]
    }
  }
  
  if (nrow(df) > 0 && df$RR[nrow(df)] > 1.3) {
    print("Removing last row with RR >", 1.3)
    df <- df[-nrow(df), ]
  }
  
  # Convert back RR to milliseconds
  df <- df %>% mutate(RR = RR * 1000)
  print("Converted RR back to milliseconds:")
  print(head(df))
  
  return(df)
}

```

### Imputation Version 3

```{r}
impute_RR_intervals <- function(df) {
  
  # library
  library(dplyr)
  
  # Convert RR and Time from milliseconds to seconds
  df <- df %>% mutate(RR = RR / 1000, Time = Time / 1000)
  
  # Helper function to compute E_10
  compute_E_10 <- function(df, i) {
    if (i + 10 > nrow(df) || i - 1 < 1) return(NA)
    sum(abs((df$RR[i:(i + 10)] - df$RR[(i - 1):(i + 9)]) / df$RR[(i - 1):(i + 9)])) / 10
  }
  
  # Helper function to compute deviation E_r and E_l
  compute_deviation <- function(RR_j, RR_prev, RR_next) {
    E_l <- abs((RR_j - RR_prev) / RR_prev)
    E_r <- abs((RR_next - RR_j) / RR_j)
    return(list(E_l = E_l, E_r = E_r))
  }
  
  # Iterate over RR intervals
  i <- 1
  while (i < nrow(df)) {
    cat("Iteration:", i, "\n")
    
    # Check if the time difference > 1.3 and not equal to RR[i+1]
    while (!is.na(df$Time[i]) && !is.na(df$Time[i + 1]) && (df$Time[i + 1] - df$Time[i]) > 1.3) {
      
      # Skip the gap if there are not enough previous RR intervals
      if (i - 9 <= 0) {
        cat("Not enough previous RR intervals, skipping gap.\n")
        break
      }
      
      # Calculate E_10 once per gap
      E_10 <- compute_E_10(df, i)
      
      attempt_count <- 0
      max_attempts <- 5  # Set a reasonable limit
      while (attempt_count < max_attempts) {
        # Use mean and std deviation from 10 previous observations
        mu <- mean(df$RR[(i - 9):i])
        sigma <- sd(df$RR[(i - 9):i])
        
        # Generate new RR value
        set.seed(999)
        new_RR <- rnorm(1, mean = mu, sd = sigma)
        cat("Generated new RR:", new_RR, "\n")
        
        # Insert the new RR value to the gap
        df <- df %>% add_row(RR = new_RR, .before = i + 1)
        
        # Insert new Time to the inserted new RR
        T_end <- df$Time[i + 2]
        new_time <- T_end - df$RR[i + 2]
        df$Time[i + 1] <- new_time
        cat("New Time value inserted:", new_time, "\n")
        
        # Check conditions for the new RR
        deviations <- compute_deviation(new_RR, df$RR[i - 1], df$RR[i + 1])
        
        if (!is.na(deviations$E_l) && !is.na(deviations$E_r) && !is.na(E_10) &&
            deviations$E_l <= E_10 && deviations$E_r <= E_10 &&
            new_RR >= 0.3 && new_RR <= 1.3) {
          # Conditions are met, continue to next gap
          break
        }
        
        # Conditions not met, remove the inserted RR and try again
        cat("Conditions not met, retrying...\n")
        df <- df[-(i + 1), ] # Remove the observation
        attempt_count <- attempt_count + 1
        sigma <- sigma * 1.05  # Loosen the requirements slightly
      }
      
      # If maximum attempts reached without success, skip to next interval
      if (attempt_count == max_attempts) {
        cat("Exceeded maximum attempts. Skipping this gap.\n")
        break
      }
    }
    i <- i + 1
  }
  
  # Convert RR and Time from seconds back to milliseconds
  df <- df %>% mutate(RR = round((RR * 1000),0), Time = round((Time * 1000),0))
  
  # Return the dataframe result
  return(df)
}
```

### Apply the imputation function to the dataset

Note: You need to change the file path with the current file path of the data you are currenly working on. 

### Participant 18682

```{r}
# input the directory
directory <- "O:/WorkLife-Data/biomonitors/holter_ecg/18682/" 
# change the directory with the current directory

# input the file path
file <- c("18682_advhrv.txt") 
# change the directory with the current directory

# file path
full_file_path <- file.path(directory, file)
  
# read the CSV file
data_original18682 <- data.table::fread(full_file_path, sep = "\t")
  
# change column name
colnames(data_original18682 )[4] <- "conv_type"
```

```{r}
# apply the helper function to one dataset
data_filtered18682<- filtering(data_original18682) 
```

```{r}
# subset the data (turned off for testing purpose)
# check for smaller dataset such as 300, 1000, 5000, 15000, ...., max observation
# extract4a <- data_filtered[1:190674, ]
```

```{r}
# apply the imputation code
data_imputed18682 <- impute_RR_intervals(data_filtered)
```

```{r}
# check the summary of filtered data
summary(data_original18682$Time)
summary(data_original18682$RR)
dim(data_original18682)
head(data_original18682)

# check the summary of data before imputation
summary(data_filtered18682$Time)
summary(data_filtered18682$RR)
dim(data_filtered18682)
head(data_filtered18682)

# check the summary of data after imputation
summary(data_imputed18682$Time)
summary(data_imputed18682$RR)
dim(data_imputed18682)
head(data_imputed18682)
```
```{r}
# save the imputed datasets
output_file <- paste0("18682_imputed.txt")

# Full output file path
output_file_path <- file.path(directory, output_file)

# Save the processed data
write.table(data_imputed18682, file = output_file_path, sep = "\t", row.names = FALSE, col.names = TRUE)
```

### Participant 49314

```{r}
# input the directory
directory <- "O:/WorkLife-Data/biomonitors/holter_ecg/49314/" 
# change the directory with the current directory

# input the file path
file <- c("49314_advhrv.txt") 
# change the directory with the current directory

# file path
full_file_path <- file.path(directory, file)
  
# read the CSV file
data_original49314 <- data.table::fread(full_file_path, sep = "\t")
  
# change column name
colnames(data_original49314)[4] <- "conv_type"
```

```{r}
# apply the helper function to one dataset
data_filtered49314 <- filtering(data_original49314) 
```

```{r}
# subset the data (turned off for testing purpose)
# check for smaller dataset such as 300, 1000, 5000, 15000, ...., max observation
# extract4a <- data_filtered[1:190674, ]
```

```{r}
# apply the imputation code
data_imputed49314 <- impute_RR_intervals(data_filtered49314)
```

```{r}
# check the summary of filtered data
summary(data_original49314$Time)
summary(data_original49314$RR)
dim(data_original49314)
head(data_original49314)

# check the summary of data before imputation
summary(data_filtered49314$Time)
summary(data_filtered49314$RR)
dim(data_filtered49314)
head(data_filtered49314)

# check the summary of data after imputation
summary(data_imputed49314$Time)
summary(data_imputed49314$RR)
dim(data_imputed49314)
head(data_imputed49314)
```

```{r}
# save the imputed datasets
output_file <- paste0("49314_imputed.txt")

# Full output file path
output_file_path <- file.path(directory, output_file)

# Save the processed data
write.table(data_imputed49314, file = output_file_path, sep = "\t", row.names = FALSE, col.names = TRUE)
```

### Participant 59503

```{r}
# input the directory
directory <- "O:/WorkLife-Data/biomonitors/holter_ecg/59503/" 
# change the directory with the current directory

# input the file path
file <- c("59503_advhrv.txt") 
# change the directory with the current directory

# file path
full_file_path <- file.path(directory, file)
  
# read the CSV file
data_original59503 <- data.table::fread(full_file_path, sep = "\t")
  
# change column name
colnames(data_original59503)[4] <- "conv_type"
```

```{r}
# apply the helper function to one dataset
data_filtered59503 <- filtering(data_original59503) 
```

```{r}
# apply the imputation code
data_imputed59503 <- impute_RR_intervals(data_filtered59503)
```

```{r}
# check the summary of filtered data
summary(data_original59503$Time)
summary(data_original59503$RR)
dim(data_original59503)
head(data_original59503)

# check the summary of data before imputation
summary(data_filtered59503$Time)
summary(data_filtered59503$RR)
dim(data_filtered59503)
head(data_filtered59503)

# check the summary of data after imputation
summary(data_imputed59503$Time)
summary(data_imputed59503$RR)
dim(data_imputed59503)
head(data_imputed59503)
```

```{r}
# save the imputed datasets
output_file <- paste0("59503_imputed.txt")

# Full output file path
output_file_path <- file.path(directory, output_file)

# Save the processed data
write.table(data_imputed59503, file = output_file_path, sep = "\t", row.names = FALSE, col.names = TRUE)
```


### Participant 85321

```{r}
# input the directory
directory <- "O:/WorkLife-Data/biomonitors/holter_ecg/85321/" 
# change the directory with the current directory

# input the file path
file <- c("85321_advhrv.txt") 
# change the directory with the current directory

# file path
full_file_path <- file.path(directory, file)
  
# read the CSV file
data_original85321 <- data.table::fread(full_file_path, sep = "\t")
  
# change column name
colnames(data_original85321)[4] <- "conv_type"
```

```{r}
# apply the helper function to one dataset
data_filtered85321 <- filtering(data_original85321) 
```

```{r}
# apply the imputation code
data_imputed85321 <- impute_rr_intervals(data_filtered85321)
```

```{r}
# check the summary of filtered data
summary(data_original85321$Time)
summary(data_original85321$RR)
dim(data_original85321)
head(data_original85321)

# check the summary of data before imputation
summary(data_filtered85321$Time)
summary(data_filtered85321$RR)
dim(data_filtered85321)
head(data_filtered85321)

# check the summary of data after imputation
summary(data_imputed85321$Time)
summary(data_imputed85321$RR)
dim(data_imputed85321)
head(data_imputed85321)
```

```{r}
# save the imputed datasets
output_file <- paste0("85321_imputed.txt")

# Full output file path
output_file_path <- file.path(directory, output_file)

# Save the processed data
write.table(data_imputed85321, file = output_file_path, sep = "\t", row.names = FALSE, col.names = TRUE)
```

### Participant 86760

```{r}
# input the directory
directory <- "O:/WorkLife-Data/biomonitors/holter_ecg/86760/" 
# change the directory with the current directory

# input the file path
file <- c("86760_advhrv.txt") 
# change the directory with the current directory

# file path
full_file_path <- file.path(directory, file)
  
# read the CSV file
data_original86760 <- data.table::fread(full_file_path, sep = "\t")
  
# change column name
colnames(data_original86760)[4] <- "conv_type"
```

```{r}
# apply the helper function to one dataset
data_filtered86760 <- filtering(data_original86760) 
```

```{r}
# apply the imputation code
data_imputed86760 <- impute_rr_intervals(data_filtered86760)
```

```{r}
# check the summary of filtered data
summary(data_original86760$Time)
summary(data_original86760$RR)
dim(data_original86760)
head(data_original86760)

# check the summary of data before imputation
summary(data_filtered86760$Time)
summary(data_filtered86760$RR)
dim(data_filtered86760)
head(data_filtered86760)

# check the summary of data after imputation
summary(data_imputed86760$Time)
summary(data_imputed86760$RR)
dim(data_imputed86760)
head(data_imputed86760)
```

```{r}
# save the imputed datasets
output_file <- paste0("86760_imputed.txt")

# Full output file path
output_file_path <- file.path(directory, output_file)

# Save the processed data
write.table(data_imputed86760, file = output_file_path, sep = "\t", row.names = FALSE, col.names = TRUE)
```

### Participant 43465

```{r}
# input the directory
directory <- "O:/WorkLife-Data/biomonitors/holter_ecg/43465/" 
# change the directory with the current directory

# input the file path
file <- c("43465_advhrv.txt") 
# change the directory with the current directory

# file path
full_file_path <- file.path(directory, file)
  
# read the CSV file
data_original43465 <- data.table::fread(full_file_path, sep = "\t")
  
# change column name
colnames(data_original43465)[4] <- "conv_type"
```

```{r}
# apply the helper function to one dataset
data_filtered43465 <- filtering(data_original43465) 
```

```{r}
# apply the imputation code
data_imputed43465 <- impute_rr_intervals(data_filtered43465)
```

```{r}
# check the summary of filtered data
summary(data_original43465$Time)
summary(data_original43465$RR)
dim(data_original43465)
head(data_original43465)

# check the summary of data before imputation
summary(data_filtered43465$Time)
summary(data_filtered43465$RR)
dim(data_filtered43465)
head(data_filtered43465)

# check the summary of data after imputation
summary(data_imputed43465$Time)
summary(data_imputed43465$RR)
dim(data_imputed43465)
head(data_imputed43465)
```

```{r}
# save the imputed datasets
output_file <- paste0("43465_imputed.txt")

# Full output file path
output_file_path <- file.path(directory, output_file)

# Save the processed data
write.table(data_imputed43465, file = output_file_path, sep = "\t", row.names = FALSE, col.names = TRUE)
```

### Participant 19003

```{r}
# input the directory
directory <- "O:/WorkLife-Data/biomonitors/holter_ecg/19003/" 
# change the directory with the current directory

# input the file path
file <- c("19003_advhrv.txt") 
# change the directory with the current directory

# file path
full_file_path <- file.path(directory, file)
  
# read the CSV file
data_original19003 <- data.table::fread(full_file_path, sep = "\t")
  
# change column name
colnames(data_original19003)[4] <- "conv_type"
```

```{r}
# apply the helper function to one dataset
data_filtered19003 <- filtering(data_original19003) 
```

```{r}
# apply the imputation code
data_imputed19003 <- impute_rr_intervals(data_filtered19003)
```

```{r}
# check the summary of filtered data
summary(data_original19003$Time)
summary(data_original19003$RR)
dim(data_original19003)
head(data_original19003)

# check the summary of data before imputation
summary(data_filtered19003$Time)
summary(data_filtered19003$RR)
dim(data_filtered19003)
head(data_filtered19003)

# check the summary of data after imputation
summary(data_imputed19003$Time)
summary(data_imputed19003$RR)
dim(data_imputed19003)
head(data_imputed19003)
```

```{r}
# save the imputed datasets
output_file <- paste0("19003_imputed.txt")

# Full output file path
output_file_path <- file.path(directory, output_file)

# Save the processed data
write.table(data_imputed19003, file = output_file_path, sep = "\t", row.names = FALSE, col.names = TRUE)
```

### Code Archive

This archive is saved here for recalling purposes (if required).

```{r}
# library(dplyr)
# 
# filtering <- function(df) {
#   
#   # Convert RR to seconds
#   df <- df %>% mutate(RR = RR / 1000)
#   
#   # Helper function to compute E_10
#   compute_E_10 <- function(df, i) {
#     if (i + 10 > nrow(df) || i - 1 < 1) return(NA)
#     sum(abs((df$RR[i:(i + 10)] - df$RR[(i - 1):(i + 9)]) / df$RR[(i - 1):(i + 9)])) / 10
#   }
#   
#   # Process the data
#   # set the starting point
#   i <- 1
#   while (i < nrow(df)) {
#     
#     # If RR > 1.3 s, then remove RR_i and its timestamp
#     # this is the new condition
#     if (!is.na(df$RR[i]) && df$RR[i] > 1.3) {
#       df <- df[-i, ]
#       next
#     }
#     
#     # if RR < 0.3s, then do all the merging process
#     if (!is.na(df$RR[i]) && df$RR[i] < 0.3) {
#       if (i + 1 > nrow(df)) break
#       RR_r <- df$RR[i] + df$RR[i + 1]
#       ER_r <- abs((RR_r - df$RR[i]) / df$RR[i])
#       ER_l <- abs((RR_r - df$RR[i - 1]) / df$RR[i - 1])
#       E_tot_r <- if (!is.na(ER_r) && !is.na(ER_l)) ER_r + ER_l else NA
#       E_10 <- compute_E_10(df, i)
#       
#       # If RR_r < 1.3s and ER_l <= E_10 and ER_r <= E_10 then 
#       # do right merge, replace RR_i+1 by RR_r and delete RR_i and its timestamp
#       if (!is.na(RR_r) && !is.na(ER_r) && !is.na(ER_l) && !is.na(E_10) && RR_r < 1.3 && ER_l <= E_10 && ER_r <= E_10) {
#         df$RR[i + 1] <- RR_r
#         df <- df[-i, ]
#         next
#       } else {
#         if (i - 1 < 1) break
#         RR_l <- df$RR[i] + df$RR[i - 1]
#         EL_r <- abs((RR_l - df$RR[i]) / df$RR[i])
#         EL_l <- abs((RR_l - df$RR[i - 1]) / df$RR[i - 1])
#         E_tot_l <- if (!is.na(EL_r) && !is.na(EL_l)) EL_r + EL_l else NA
#         E_10_prev <- compute_E_10(df, i - 1)
#         
#         # If RR_l < 1.3 s and EL_l <= E10 and EL_r <= E_10 then 
#         # left merge, replace RR_i by RR_l and delete RR_i-1 and its timestamp
#         if (!is.na(RR_l) && !is.na(EL_l) && !is.na(EL_r) && !is.na(E_10_prev) && RR_l < 1.3 && EL_l <= E_10_prev && EL_r <= E_10_prev) {
#           df$RR[i - 1] <- RR_l
#           df <- df[-i, ]
#           next
#         } else if (!is.na(RR_r) && !is.na(RR_l) && RR_r > 1.3 && RR_l > 1.3) {
#           if (i + 1 <= nrow(df)) { # Else if RR_r > 1.3 and RR_l > 1.3, then Delete both {RR}_i and {RR}_{i+1}
#             df <- df[-c(i, i + 1), ]
#           } else {
#             df <- df[-i, ]
#           }
#           next
#         } 
#         # Else if RR_r < 1.3 and RR_l > 1.3 then Replace {RR}_{i+1} by {RR}_r and delete {RR}_i and its timestamp
#         else if (!is.na(RR_r) && !is.na(RR_l) && RR_r < 1.3 && RR_l > 1.3) {
#           df$RR[i + 1] <- RR_r 
#           df <- df[-i, ]
#           next
#         } 
#         # Else if RR_r > 1.3s and RR_l < 1.3s then Replace {RR}_{i-1} by {RR}_l and delete {RR}_i and its timestamp
#         else if (!is.na(RR_r) && !is.na(RR_l) && RR_r > 1.3 && RR_l < 1.3) {
#           df$RR[i - 1] <- RR_l 
#           df <- df[-i, ]
#           next
#         } 
#         # Else if both {RR}_l and {RR}_r<\ 1.3\ s but errors {Etot}_r and {Etot}_l are higher than 0.4, then keep the one with smaller error
#         else if (!is.na(RR_l) && !is.na(RR_r) && !is.na(E_tot_r) && !is.na(E_tot_l) && RR_l < 1.3 && RR_r < 1.3 && E_tot_r > 0.4 
#                  && E_tot_l > 0.4) {
#           if (E_tot_r < E_tot_l) { 
#             df$RR[i + 1] <- RR_r
#             df <- df[-i, ]
#           } else {
#             df$RR[i - 1] <- RR_l
#             df <- df[-i, ]
#           }
#           next
#         }
#       }
#     }
#     # restate the iteration
#     i <- i + 1
#   }
#   
#   # Handle the last row
#   # if the RR < 0.3
#   while (nrow(df) > 0 && df$RR[nrow(df)] < 0.3) {
#     
#     # compute RR left
#     RR_l <- df$RR[nrow(df)] + df$RR[nrow(df)-1]
#     
#     # if RR_l > 1.3 s
#     if (RR_l > 1.3) {
#       df <- df[-nrow(df), ] # remove the row
#       next
#     } else {
#       df$RR[nrow(df)-1] <- RR_l # otherwise, do left merge
#       df <- df[-nrow(df), ]
#     }
#   }
#   
#   # if RR > 1.3
#   if (nrow(df) > 0 && df$RR[nrow(df)] > 1.3) {
#     df <- df[-nrow(df), ] # remove the row
#   }
#   
#   # Convert back RR to milliseconds
#   df <- df %>% mutate(RR = RR * 1000)
#   
#   # Return the dataframe result
#   return(df)
# }
```

### Imputation Code

```{r}
# library(dplyr)
# 
# impute_rr_intervals <- function(df) {
#   
#   # Convert RR and Time from milliseconds to seconds
#   df <- df %>% mutate(RR = RR / 1000, Time = Time / 1000)
#   
#   # Helper function to compute E_10
#   compute_E_10 <- function(df, i) {
#     if (i + 10 > nrow(df) || i - 1 < 1) return(NA)
#     sum(abs((df$RR[i:(i + 10)] - df$RR[(i - 1):(i + 9)]) / df$RR[(i - 1):(i + 9)])) / 10
#   }
#   
#   # Helper function to compute deviation E_r and E_l
#   compute_deviation <- function(RR_j, RR_prev, RR_next) {
#     E_l <- abs((RR_j - RR_prev) / RR_prev)
#     E_r <- abs((RR_next - RR_j) / RR_j)
#     return(list(E_l = E_l, E_r = E_r))
#   }
#   
#   # Iterate over RR intervals
#   i <- 1
#   while (i < nrow(df)) {
#     cat("Iteration:", i, "\n")
#     
#     # Check if the time difference > 1.3 and not equal to RR[i+1]
#     while (!is.na(df$Time[i]) && !is.na(df$Time[i + 1]) && (df$Time[i + 1] - df$Time[i]) > 1.3) {
#       
#       # Skip the gap if there are not enough previous RR intervals
#       if (i - 9 <= 0) {
#         cat("Not enough previous RR intervals, skipping gap.\n")
#         break
#       }
#       
#       attempt_count <- 0
#       inserted_data_count <- 0
#       repeat {
#         # Use mean and std deviation from 10 previous observations
#         mu <- mean(df$RR[(i - 9):i])
#         sigma <- sd(df$RR[(i - 9):i])
#         
#         # Generate new RR value
#         # set.seed(999)
#         new_rr <- rnorm(1, mean = mu, sd = sigma)
#         cat("Generated new RR:", new_rr, "\n")
#         
#         # Insert the new RR value to the gap
#         df <- df %>% add_row(RR = new_rr, .before = i + 1)
#         inserted_data_count <- inserted_data_count + 1
#         
#         # Insert new Time to the inserted new RR
#         T_end <- df$Time[i + 2]
#         new_time <- T_end - df$RR[i + 2]
#         df$Time[i + 1] <- new_time
#         cat("New Time value inserted:", new_time, "\n")
#         
#         # Check conditions for the new RR
#         deviations <- compute_deviation(new_rr, df$RR[i - 1], df$RR[i + 1])
#         E_10 <- compute_E_10(df, i)
#         
#         if (!is.na(deviations$E_l) && !is.na(deviations$E_r) && !is.na(E_10) &&
#             deviations$E_l <= E_10 && deviations$E_r <= E_10 && new_rr >= 0.3 && new_rr <= 1.3) {
#           # Conditions are met, continue to next gap
#           break
#         }
#         
#         # Conditions not met, retry
#         attempt_count <- attempt_count + 1
#         if (attempt_count > 3) {
#           cat("Exceeded maximum attempts.\n")
#           if (inserted_data_count == 1) {
#             df <- df[-(i + 1), ] # Remove one observation
#           } else {
#             df <- df[-((i + 1):(i + 2)), ] # Remove two observations
#             for (j in 1:4) {
#               E_10 <- E_10 * 1.05 # Loosen the requirements by 5%
#             }
#           }
#           break
#         }
#         
#         cat("Conditions not met, retrying...\n")
#         df <- df[-(i + 1), ] # Remove observation
#         inserted_data_count <- inserted_data_count - 1
#         E_10 <- E_10 * 1.05 # Loosen the requirements by 5%
#       }
#     }
#     i <- i + 1
#   }
#   
#   # Convert RR and Time from seconds back to milliseconds
#   df <- df %>% mutate(RR = RR * 1000, Time = Time * 1000)
#   
#   # Return the dataframe result
#   return(df)
# }
```

### Imputation version 2

```{r}
# library(dplyr)
# 
# impute_rr_intervals <- function(df, max_attempts = 3, sigma_loosen = 1.05) {
#   
#   # Convert RR and Time from milliseconds to seconds
#   df <- df %>% mutate(RR = RR / 1000, Time = Time / 1000)
#   
#   # Helper function to compute E_10
#   compute_E_10 <- function(df, i) {
#     if (i + 10 > nrow(df) || i - 1 < 1) return(NA)
#     sum(abs((df$RR[i:(i + 10)] - df$RR[(i - 1):(i + 9)]) / df$RR[(i - 1):(i + 9)])) / 10
#   }
#   
#   # Helper function to compute deviation E_r and E_l
#   compute_deviation <- function(RR_j, RR_prev, RR_next) {
#     E_l <- abs((RR_j - RR_prev) / RR_prev)
#     E_r <- abs((RR_next - RR_j) / RR_j)
#     return(list(E_l = E_l, E_r = E_r))
#   }
#   
#   # Iterate over RR intervals
#   i <- 1
#   while (i < nrow(df)) {
#     cat("Iteration:", i, "\n")
#     
#     # Check if the time difference > 1.3 and not equal to RR[i+1]
#     while (!is.na(df$Time[i]) && !is.na(df$Time[i + 1]) && (df$Time[i + 1] - df$Time[i]) > 1.3) {
#       
#       # Skip the gap if there are not enough previous RR intervals
#       if (i - 9 <= 0) {
#         cat("Not enough previous RR intervals, skipping gap.\n")
#         break
#       }
#       
#       # Compute E_10 once before retry loop
#       E_10 <- compute_E_10(df, i)
#       
#       attempt_count <- 0
#       inserted_data_count <- 0
#       
#       repeat {
#         # Use mean and std deviation from 10 previous observations
#         mu <- mean(df$RR[(i - 9):i])
#         sigma <- sd(df$RR[(i - 9):i]) * sigma_loosen^attempt_count # Loosen sigma with each attempt
#         
#         # Generate new RR value
#         #set.seed(999)
#         new_rr <- rnorm(1, mean = mu, sd = sigma)
#         cat("Generated new RR:", new_rr, "\n")
#         
#         # Insert the new RR value to the gap
#         df <- df %>% add_row(RR = new_rr, .before = i + 1)
#         inserted_data_count <- inserted_data_count + 1
#         
#         # Insert new Time corresponding to the inserted new RR
#         T_end <- df$Time[i + 2]
#         new_time <- T_end - df$RR[i + 2]
#         df$Time[i + 1] <- new_time
#         cat("New Time value inserted:", new_time, "\n")
#         
#         # Check conditions for the new RR
#         deviations <- compute_deviation(new_rr, df$RR[i - 1], df$RR[i + 1])
#         
#         if (!is.na(deviations$E_l) && !is.na(deviations$E_r) &&
#             deviations$E_l <= E_10 && deviations$E_r <= E_10 &&
#             new_rr >= 0.3 && new_rr <= 1.3) {
#           # Conditions are met, continue to the next gap
#           break
#         }
#         
#         # Conditions not met, retry
#         attempt_count <- attempt_count + 1
#         
#         if (attempt_count >= max_attempts) {
#           cat("Exceeded maximum attempts.\n")
#           if (inserted_data_count == 1) {
#             df <- df[-(i + 1), ] # Remove one observation
#           } else {
#             df <- df[-((i + 1):(i + 2)), ] # Remove two observations
#           }
#           break
#         }
#         
#         # Remove the inserted observation and retry
#         df <- df[-(i + 1), ] # Remove the last inserted RR
#         inserted_data_count <- inserted_data_count - 1
#         cat("Retrying with loosened sigma...\n")
#       }
#     }
#     i <- i + 1
#   }
#   
#   # Convert RR and Time from seconds back to milliseconds
#   df <- df %>% mutate(RR = RR * 1000, Time = Time * 1000)
#   
#   # Return the modified dataframe
#   return(df)
# }
```