---
title: "Statistics_Group"
author: "Aulia Dini Rafsanjani"
date: "`r Sys.Date()`"
output: html_document
---

### Goal

The goal of this document is to provide the summary statistics as a part of statistics group discussion. We provided the descriptive statistics of heart rate data that we got from participants. The descriptive statistics includes :

1. Data dimension, duration of wearing device in days and hours. 

2. Summary statistics of RR

3. Percentage of flags provided by the device. 

4. Plot from each participant. 

### Data structure

The data has two variable : Time and R-R interval. 

The data covered nine participant. 

The data covered observation from seven days recording. Some participants have a break-off because of irritation and device issue. 

### Link to statistics group manuscript

https://docs.google.com/document/d/1Q1r4xX8rZGfGTwUjeTDKml9DDDBJ7Vi8/edit?usp=sharing&ouid=111075353933698324521&rtpof=true&sd=true

### Library

```{r}
library(data.table)
library(dplyr)
library(tidyr)
library(data.table)
library(dplyr)
library(ggplot2)
library(gridExtra)
```

### Overall summary statistics

```{r}
# Define directory and file list
directory <- "/Users/auliadinirafsanjani/Dropbox (University of Michigan)/WorkLife_ECGTest/Test"

file_list <- c(
  "testben1_ADVHRV_2024-02-02_1244.txt",
  "testdini1_ADVHRV_20240112.txt", 
  "testamanda1_ADVHRV_2024-01-26.txt",
  "testjenny1_ADVHRV_2024-01-12.txt",
  "testkaris1_ADVHRV_2024-01-26.txt",
  "testmary1_ADVHRV_2024-01-12.txt",
  "testneil1_ADVHRV_2024-01-15.txt",
  "test190031_ADVHRV_2024-05-07.txt", 
  "test867601_ADVHRV_2024-05-07.txt"
)

result_df1 <- data.frame()
result_df2 <- data.frame()
result_df3 <- data.frame()
plots_list <- list()

for (file in file_list) {
  
  # File path
  full_file_path <- file.path(directory, file)
  
  # Read the CSV file
  data <- fread(full_file_path, sep = "\t")
  
  # Change column name
  colnames(data)[4] <- "category"
  
  # Create table of category percentages
  table1 <- data %>%
    summarise(count = n(), 
              time_days = round((max(Time)/(1000*86400)),1), # duration in days
              time_hours = round((max(Time)/(1000*3600)),1)) %>% # duration in hours
    mutate(data_name = gsub("\\..*$", "", file))
  
  
  # Create table of RR summary (in seconds)
  table2 <- data %>%
    summarise(minimum = round((min(RR)/1000),2),
              q1 = round((quantile(RR, 0.25)/1000),2),
              median = round((median(RR)/1000),2),
              q3 = round((quantile(RR, 0.75)/1000),2),
              maximum = round((max(RR)/1000),2), 
              mean = round((mean(RR)/1000),2), 
              sd = round((((sd(RR))/1000)),2)) %>%
    mutate(data_name = gsub("\\..*$", "", file))
  
  # Create table of category percentages # interest is in category U and I
  table3 <- data %>%
    group_by(category) %>%
    summarize(percentage =round((n() / nrow(data) * 100),3)) %>%
    mutate(data_name = gsub("\\..*$", "", file))
  
  # Create a line plot for each file
  plot <- ggplot(data = data, aes(x = Time, y = RR)) +
    geom_line() +  
    labs(x = "Time", y = "RR") +  # Label x and y axes
    ggtitle(paste("Heart Rate vs. Time -", gsub("\\..*$", "", file))) +  
    theme_minimal()  
  
  # Append the percentage table to the result data frame
  result_df1 <- bind_rows(result_df1, table1)
  result_df2 <- bind_rows(result_df2, table2)
  result_df3 <- bind_rows(result_df3, table3)
  
  # Store the plot in the list
  plots_list[[file]] <- plot
}

# Reorder columns if needed
# Print Result
result_df1 <- result_df1[, c("data_name", "count", "time_days", "time_hours")]
result_df1

result_df2 <- result_df2[, c("data_name", "minimum", "q1", "median", "q3", "maximum", "mean", "sd")]
result_df2

result_df3 <- result_df3[, c("data_name", "category", "percentage")]
wide_result_df3 <- pivot_wider(result_df3, names_from = category, values_from = percentage)
colnames(wide_result_df3)[2:ncol(wide_result_df3)] <- paste0("Percent ", colnames(wide_result_df3)[2:ncol(wide_result_df3)])
wide_result_df3

plots_list
```
```{r}
# change the data name
# result_df <- result_df %>%
#   mutate(data_id = case_when(
#     data_name == "testben1_ADVHRV_2024-02-02_1244" ~ "1",
#     data_name == "testdini1_ADVHRV_20240112" ~ "2",
#     data_name == "testamanda1_ADVHRV_2024-01-26" ~ "3",
#     data_name == "testjenny1_ADVHRV_2024-01-12" ~ "4",
#     data_name == "testkaris1_ADVHRV_2024-01-26" ~ "5",
#     data_name == "testmary1_ADVHRV_2024-01-12" ~ "6",
#     data_name == "testneil1_ADVHRV_2024-01-15" ~ "7",
#     data_name == "test190031_ADVHRV_2024-05-07" ~ "8",
#     data_name == "test867601_ADVHRV_2024-05-07" ~ "9",
#     TRUE ~ data_name  
#   ))
# 
# head(result_df)
# 
# result_df <- result_df[, c("data_name", "data_id", "minimum", "q1", "median", "q3", "maximum", "mean", "sd")]
# 
# print(result_df[,2:9])
```

```{r}
# Save plots as JPEG files
# for (file_name in names(plots_list)) {
#   jpeg(file.path(directory, paste0(gsub("\\..*$", "", file_name), ".jpeg")))
#   print(plots_list[[file_name]])
#   dev.off()
# }
```

### Further work :

1. Check the naming format in the O: drive. 

2. Create code to access the data based on each folder name and file name. 
